### 2015-05-16 manuscript/manuscript_cordiv.tex
Initial commit of files

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{nicefrac}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{footnote}
\usepackage{rotating}
%\usepackage{slashbox}
\usepackage{afterpage}
\usepackage{float}
\usepackage{color}

\usepackage[margin = 1.0 in]{geometry}
\usepackage{natbib}

\renewcommand{\bottomfraction}{.9}
\renewcommand{\topfraction}{.9}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{.9}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   document style macros
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\gtrsim{\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}}
\def\lessim{\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}}
\newcommand{\ddg}{$\Delta\Delta G~$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\graphicspath{{../../analysis/figures/}}

\title{Sequence divergence as the main determinant of sequence-structure relationships}
\author{Amir Shahmoradi$^{1*}$, Eleisha L. Jackson$^2$, Claus O. Wilke$^2$}
\begin{document}

\date{\today}
\maketitle

\noindent
$^1$ Department of Physics, The University of Texas at Austin, Austin, TX 78712, USA \\
$^2$ Institute of Cellular and Molecular Biology, Center for Computational Biology and Bioinformatics, and Department of Integrative Biology, The University of Texas at Austin, Austin, Texas, 78712 USA\\

\bigskip
\noindent
$^*$Corresponding author\\
$\phantom{^*}$Email: amir@physics.utexas.edu\\
%$\phantom{^*}$Phone:{ \color{red} Need a phone} \\
$\phantom{^*}$Phone: +1 512 232 2459\\

\bigskip
\noindent
Manuscript type: research article\\
\bigskip
\noindent  Keywords: protein evolution, relative solvent accessibility, site variability

\begin{abstract}
NEED TO EDIT THE ABSTRACT\\
Recent work has shown that structural properties are capable of predicting site-specific sequence variability for a given protein. However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.  Here we present the results from a comprehensive search for potential biophysical and structural determinants of protein evolution by studying more than $200$ structural and evolutionary properties in a dataset of $209$ monomeric enzymes. We discuss the main protein characteristics responsible for the general patterns of protein evolution, and identify sequence divergence as the main determinant of the strengths of virtually all structural-evolution relationships, explaining $10-30 \%$ of observed variation in sequence-structure relations. In addition to sequence divergence, we identify several protein structural properties that are moderately but significantly coupled with the strength of sequence-structure relations. In particular, proteins with more homogeneous back-bone hydrogen bond energies, large fractions of helical secondary structures and low fraction of beta sheets tend to have the strongest correlations between structural properties and site variability.

\end{abstract}
\vfill
\vfill
\def\thefootnote{\fnsymbol{footnote}}
\setcounter{footnote}{0}

\section{Introduction}
\label{sec:intro}

Proteins are subject to a number of biophysical and functional constraints \citep{Scherreretal2012, Wilkeetal2010}. These constraints result site-specific patterns of sequence variability within a protein. Recently several site-specific structural properties that can explain patterns of sequence variability in proteins have been identified. One of the earliest examples is Relative Site Accessibility. \cite{Fransozaetal2009} identified RSA as the strongest predictor of evolutionary rate. They found that residues that are buried in the core of proteins tend to be more conserved than exposed residues close to the surface of the protein. In their analysis these concerned both RSA and various definitions of residue packing density to predict evolutionary rate. They found that RSA and evolutionary rate shared a significant linear relationship. Afterwards, several other works \citep{Ramseyetal2011, Scherreretal2012} also found that RSA as a very significant predictor of evolutionary rate and found this linear relationship as well. However in these papers all have the same flaw. During the course of their analysis they binned the protein sites and average over all sites within a bin when looking at the trend of RSA. This could produce artifacts when looking at the trend. \\
\indent Later, Yeh et al performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by CN and WCN \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability. A year later, Sharamoradi et al also performed a site-wise analysis on a series of viral proteins. In this analysis they found that RSA had the strongest correlation with site variability. Additionally the effect seen between CN and WCN was of a much smaller magnitude as compared to \citep{Yehetal2014}. Here we attempt to reconcile the work done in this area. We find that site variability is the primary determinant of structure determinant of the strength of structure-sequence relationships and some differences in previous work can be explained in terms of differing levels of site variability. \\
COPY EDIT THIS!!! \\

\section{Materials and Methods}
\label{sec:mam}

    \subsection*{Structures, sequences, and measures of sequence properties } 
    The results presented in this work are based on a two datasets. The first is a dataset of $209$ monomeric enzymes from \citep{Echaveetal2015} originally from Huang et al. (2014).The original dataset was comprised of $213$ but we removed four of the proteins (1BBS, 1BS0, 1DIN, 2HPL) that had did not have data at insertion sites. Briefly, these proteins are all enzyme monomers   randomly picked from the Catalytic Site Atlas $2.2.11$ \citep{Porteretal2004} with protein sizes in the sample ranging from $95$ to $1287$. For each structure we had a corresponding alignment of up to 300 homologous sequences.  The second dataset was from taken from Sharamoradi et al. and is comprised of nine? viral proteins. The viral proteins range from 122 - 557 in length and each structure is accompanied by a sequence alignment of up to 2362 homologous sequences. Sequence alignments for both datasets constructed using amino-acid sequences with MAFFT \citep{Katohetal2002, Katohetal2005}, specifying the auto flag to select the optimal algorithm for the given data set.The alignments were then used to calculate the site-specific evolutionary rates for each individual protein in both datasets. To do so, we relied on two independent methods of measuring sequence variability measure. First, we calculated the Shannon entropy ($H_i$) -- the sequence entropy, hereafter abbreviated as {\it seqent} -- at each alignment column $i$:
    \begin{equation}
        \label{eqn:shannon}
        H_i = -\sum_j P_{ij}\ln P_{ij}
    \end{equation}

    where $P_{ij}$ is the relative frequency of amino acid $j$ at position $i$ in the alignment. The sequence entropy is a measure of variability at each site. We also calculated a measure of site-specific evolutionary rate -- hereafter abbreviated as {\it r4s} -- for each protein using software Rate4site. First the Maximum Likelihood phylogenetic trees were inferred with RAxML, using the LG substitution matrix and the CAT model of rate heterogeneity \citep{Stamatakis2006, Stamatakis2014}. For each structure, we then used the respective sequence alignment and phylogenetic tree to infer site-specific substitution rates with Rate4Site, using the empirical Bayesian method and the amino-acid Jukes-Cantor mutational model (aaJC) \citep{Mayroseetal2004}.

    \subsection*{Calculation of Structural Properties}

    HOW IS CN and WCN Defined and Calculated??? (Cite Shih et al for WCN) Possbily Lin Paper \\
\indent We used DSSP \citep{Kabschetal2005}  for the calculation of Accessible Surface Area (ASA) for each site. We normalized the ASA for each site by the theoretical maximum solvent accessibility values of \cite{Tienetal2013}  to obtain the Relative Solvent Accessibility (RSA) for all individual sites in all proteins. In addition to ASA values, we also extract from DSSP output, information about the secondary structure of proteins such as the total number of residues participating in different types of helices, parallel or anti-parallel beta sheets, or loops and turns. To complete the list of pdb-level structural properties, we also calculate the Spearman correlations between all residue-level structure and sequence properties and include them in the analysis to probe their potential effects on the strength of structure-sequence relations. \\

All data and analysis scripts required to reproduce the work are publicly available to view and download at \url{https://github.com/shahmoradi/cordiv}.

\section{Results}
\label{sec:results}

    \subsection*{}
{\color{red}SOME RESULTS.....} \\

\section{Discussion}
\label{sec:dcr}
{\color{red}SOME DISCUSSION.....} \\

\section{Acknowledgements}
The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing High-performance computing resources. ELJ is is funded by a National Science Graduate Research Fellowship, grant number DGE-1110007. COW is funded by {\color{red} Which grants??}.  AS is funded by {\color{red} Which grants??}.
....

%\clearpage
%\newpage

%\cleardoublepage

\bibliographystyle{peerj} %"style
\bibliography{ej_manuscript_bib} %expected file "my refs.bib"

\section*{Figures}

1. WCN-Rate4site vs Variance of RateSite (Also for RSA) and CN-Rate4site vs Variance of RateSite \\
2. WCN-Entropy vs Mean Entropy  and CN-Entropy vs Mean Entropy \\
3. RSA-Entropy vs Mean Entropy \\
4. RSA vs Rate4Site \\
5. Should I make the stress  and delta delta G correlations plots for the viral sequences that would complete the story?

%    \begin{figure}[tbh]
%            \begin{center}
%            %\includegraphics[width=6.9in]{cordiv_similarities_r4sJC.pdf}
%            \includegraphics[width=6.9in]{cordiv_similarities_seqent.pdf}
%            \end{center}
%            \caption{A comparison of the strength of Spearman correlation of sequence evolutionary rates (r4sJC) with {\it side chain} Weighted Contact Number vs. correlations of other structural properties with evolutionary rates. Detailed description of the structural properties is given in Section \label{sec:mam}. The red lines in each plot represent equality line. It is evident from all plots that for any given protein in dataset, the correlation strength of one structural property is a good proxy measure of the correlation strength of any other structural property with sequence variability measures. For brevity, correlations of structure-rate4site are not shown here but are available online, also in supplementary material.}
%            \label{fig:seqent_structure_cors}
%    \end{figure}
%
%
%    \begin{figure}[tbh]
%        \centering
%        \includegraphics[width=3.6in]{/main_modulators_cormat_r4s_wcn.pdf} \\
%        \caption{{\bf The Spearman correlation matrix of the strongest sequence--structure correlations and the prominent determinants of the strengths of the corresponding relations}. The variables on the diagonal elements of the matrix from top to bottom represent respectively, the four strongest sequence--structure relations: r4sJC--wcnSC, r4sJC--varea, seqent--wcnSC, seqent--varea, followed by protein properties that modulate the strength of these relations: the correlation between r4sJC--seqent ({\it r.r4sJC.seqent}), variance of r4sJC ({\it var.r4sJC}), variance of sequence entropy ({\it var.seqent}), variance of back-bone hydrogen bond energies ({\it var.hbe}), and the fraction of amino acids in helical \& $\beta-$sheet secondary structures in the proteins ({\it mn.helix} \& {\it mn.betas} respectively).}
%        \label{fig:mainmods}
%    \end{figure}

%    \begin{figure*}[tbh]
  %      \centering
 %       \includegraphics[width=3.4in]{/validation_r_r4s_wcn_var_seqent.pdf}
 %       \caption{{\bf Sequence--structure correlation strength versus sequence divergence.} The plot illustrates the relationship between the strength of a representative sequence--structure correlation ({\it seqent--scnSC}) and the sequence divergence as measured by the variance of protein sequence entropy. The black circles represent $209$ proteins used in this work. For comparison and validation, the red circles represent data for $9$ viral proteins taken from {\color{red}shahmoradi2014xx}}.
%        \label{fig:validation}
%    \end{figure*}

%    \begin{figure}[tbh]
%        \centering
%        \includegraphics[width=7in]{/best_wcn/select_variables/boxplot_wcn_all_in_one.pdf} \\
%        \caption{NEEDS A CAPTION.}
%        \label{fig:best_wcn}
%    \end{figure}
%
%    \begin{figure*}[tbh]
%        \centering
%        \begin{tabular}{cc}
%        \includegraphics[width=3.4in]{/modulators_similarity_Sr4sJC_scors.pdf} & \includegraphics[width=3.4in]{/modulators_similarity_Sseqent_scors.pdf} \\
%        \end{tabular}
%        \caption{{\bf Spearman correlation matrices illustrating the similarity of the set of structure and sequence properties that contribute to the strength of sequence--structure relations}.  {\bf Left:} The correlation matrix for the relation of evolutionary rates ({\it r4sJC}) with structural quantities on the diagonal of the matrix. {\bf Right:} The correlation matrix for the relation of sequence entropy ({\it seqent}) with structural quantities on the diagonal of the matrix.}
%        \label{fig:simcormats}
%    \end{figure*}
%
%    \begin{figure*}[tbh]
%        \centering
%        %\begin{tabular}{cc}
%        \includegraphics[width=3.4in]{/validation_r_seqent_wcn_var_seqent.pdf}
%        %\end{tabular}
%        \caption{{\bf Sequence--structure correlation strength versus sequence divergence.} The plot illustrates the relationship between the strength of a representative sequence--structure correlation ({\it seqent--scnSC}) and the sequence divergence as measured by the variance of protein sequence entropy. The black circles represent $209$ proteins used in this work. For comparison and validation, the red circles represent data for $9$ viral proteins taken from {\color{red} shahmoradi2014xx}}.
%        \label{fig:validation}
%    \end{figure*}

\end{document}

### 2015-05-16 scripts/get_data.R

@@ -1,0 +1,6 @@
+library(dplyr)
+enzyme_data = read.table(file = "../data/jec_ddg_209_monomers.csv", sep = ",", header = T)
+enzyme_out_data = data.frame(enzyme_data$pdb, enzyme_data$nsites, enzyme_data$nseq, enzyme_data$site, enzyme_data$rsa.tien, enzyme_data$r4s.jc.ej, enzyme_data$r4s.jtt.ej)
+enzyme_out_data = rename(enzyme_out_data,  pdb_id = enzyme_data.pdb, num_sites = enzyme_data.nsites, num_seqs = enzyme_data.nseq, site =  enzyme_data.site, rsa = enzyme_data.rsa.tien, r4c_jc = enzyme_data.r4s.jc.ej, r4s.jtt = enzyme_data.r4s.jtt.ej)
+enzyme_out_data = mutate(enzyme_out_data, data_id = "enzyme")
+write.csv(enzyme_out_data, "../data/huang_data.csv")

### 2015-05-19 entropies/scripts/calculate_sequence_entropy.py
Adding files... Need to add readmes and 1RD8_AB files

@@ -1,0 +1,95 @@
+#!/usr/bin/python
+import protein_toolbox_helper as ph
+import numpy as np
+import subprocess
+import sys
+
+'''
+Last Updated: November 21, 2014
+Description: This is a script that calculates entropy values for an alignment
+'''
+
+aa_list = ['G', 'A', 'L', 'V', 'I', 'P', 'R', 'T', 'S', 'C', 'M', 'K', 'E', 'Q', 'D', 'N', 'W', 'Y', 'F', 'H'] #List of amino acids
+
+def calculate_aa_frequencies(list_of_aligned_seqs): 
+	'''
+	This function calculates the amino acid frequencies for each site in an alignment
+	Args: 
+		list_of_aligned_seqs: A list of sequences that have been aligned
+	
+	Returns:
+		all_freqs: A list of lists where each list are the amino acid frequencies for a site
+	
+	'''
+	all_freqs = []
+	aligned_seqs = []
+	for seq in list_of_aligned_seqs:
+		aligned_seqs.append(ph.get_sequence_array(seq)) #Turn the lists of sequences into a list of arrays (where each array represents a sequence 
+	seq_alignment = np.array(aligned_seqs)
+	rows, cols = seq_alignment.shape
+
+	for i in xrange(0, cols): #For each column in the alignment numpy array (each site in the alignment)
+		#print "site: ", i
+		site_freqs = []
+		site = seq_alignment[:,i].tolist()
+		num_aas_at_site = len(site) - site.count('-') #Count the number of amino acids at the site
+		for aa in aa_list: #For each of the 20 amino acid types
+			aa_freq = float(site.count(aa))/float(num_aas_at_site) #Calculate the AA's frequences
+			site_freqs.append(aa_freq) #Append it to the list of frequencies for that site
+		all_freqs.append(site_freqs) #Append it to the list of lists for the frequencies
+	return all_freqs
+
+def calculate_entropy(data_array):
+	'''
+	Calculates the entropy of list of lists where each list represents a site
+	
+	Args:
+		data_array: A 2D numpy array where the rows are sites and the columns represent a value for each of the 20 amino acids
+			(The values could be its frequency at that site or another meaningful calculated quantity)
+	Returns:
+		entropy_values: The entropy value for each site	
+	'''
+	
+	entropy_values = []
+	num_residues,num_AA = data_array.shape  
[…]

### 2015-05-19 entropies/scripts/create_pdb_to_alignment_map.py

@@ -1,0 +1,78 @@
+#!/usr/bin/python
+import protein_toolbox_helper as ph
+import subprocess, os, sys, re
+
+'''
+Last Edited By: Eleisha Jackson (February 2, 2015)
+Description: This script create a map. This map maps a position in the alignment to each position in the pdb sequence. Must have MAFFT installed to use this script.
+'''
+
+def create_pdb_alignment_map(seqs, pdb_res_nums, map_filename):
+	'''
+	Creates a file that maps a pdb seqeunce onto a given alignments. 
+	Args: 
+		seqs: A list of aligned sequences (
+		pdb_res_nums: A list of residue numbers from the pdb file
+		map_filename: The file name to write the map to
+	
+	Returns:
+		Does not return anything but creates a file with a map
+	'''
+	
+	alignment_length = len(seqs[0]) 
+	aligned_pdb_seq = seqs[0]
+	#Make map list
+	map_outfile = open(map_filename, "w")
+	map_outfile.write("align_pos\tpdb_pos\tpdb_aa\n")
+	res_counter = 0
+	for j in xrange(0, alignment_length):
+		map_outfile.write(str(j+1) + "\t")
+		if(aligned_pdb_seq[j]!= "-"):
+			map_outfile.write(str(pdb_res_nums[res_counter]) + "\t" + aligned_pdb_seq[j])
+			res_counter = res_counter + 1
+		else:
+			map_outfile.write("NA" + "\t" + "-")
+		map_outfile.write("\n")
+
+def main():
+	if len( sys.argv ) != 4:
+		print '''      Usage:      '''
+		print "     ", sys.argv[0], "<input alignment>", "<pdb file> " "<output map file>"
+	else:
+		align_file = sys.argv[1]
+		pdb_file = sys.argv[2]
+		map_filename = sys.argv[3]
+		
+		fileparts = re.split("/", pdb_file)
+		parts = fileparts[-1]
+		pdb_parts = re.split("_", re.split("\.", parts)[0])
+		pdb = pdb_parts[0]
+		chain = pdb_parts[1]
+		
+		(pdb_seq, pdb_res_nums) = ph.get_info_from_pdb(pdb_file, chain)	#Get the residue list from pdb
+		
+		#Get unaligned sequences from the alignment
+		(all_seqs, all_headers) = ph.get_unaligned_seqs(align_file)	
+
+		#Then add this to the top of the file
+		align_outfile = "temp.txt"
+		all_seqs.insert(0, pdb_seq)
[…]

### 2015-05-19 entropies/scripts/protein_toolbox_helper.py

@@ -1,0 +1,452 @@
+#!/usr/bin/python
+import re, os, math, string, subprocess
+from Bio.PDB.PDBIO import PDBIO
+from Bio.PDB.PDBParser import PDBParser
+from Bio import AlignIO
+from Bio import PDB
+
+#This is a dictionary that relates the three letter amino acid abbreviation with its one letter abbreviation
+resdict = { 'ALA': 'A', 'CYS': 'C', 'ASP': 'D', 'GLU': 'E', 'PHE': 'F', \
+            'GLY': 'G', 'HIS': 'H', 'ILE': 'I', 'LYS': 'K', 'LEU': 'L', \
+            'MET': 'M', 'ASN': 'N', 'PRO': 'P', 'GLN': 'Q', 'ARG': 'R', \
+            'SER': 'S', 'THR': 'T', 'VAL': 'V', 'TRP': 'W', 'TYR': 'Y' }
+
+_hydrogen=re.compile("[123 ]*H.*") 
+
+class ChainSelector(object): 
+	"""
+	Adapted from the extract() function of Biopython 
+	"""
+
+	def __init__(self, chain_id, start, end, model_id=0): 
+		self.chain_id=chain_id 
+		self.start=start 
+		self.end=end 
+		self.model_id=0 
+
+	def accept_model(self, model): 
+		# model - only keep model 0 
+		if model.get_id()==self.model_id: 
+			return 1 
+		return 0 
+
+	def accept_chain(self, chain): 
+		if chain.get_id()==self.chain_id: 
+			return 1 
+		return 0 
+		
+	def accept_residue(self, residue): 
+		# residue - between start and end 
+		hetatm_flag, resseq, icode=residue.get_id() 
+		if hetatm_flag!=" ": 
+			# skip HETATMS 
+			return 0 
+		if self.start<=resseq<=self.end: 
+			return 1 
+			return 0 
+
+	def accept_atom(self, atom): 
+		# atoms - get rid of hydrogens 
+		name=atom.get_id() 
+		if _hydrogen.match(name): 
+			return 0 
+		else: 
+			return 1 
+
+def extract(structure, chain_id, start, end, filename): 
+	""" 
+	Write out selected residues from a particular given pdb structure to filename. 
+	
+	Args: 
+		structure: The four letter pdb code for the protein (this pdb must currently exist!)
+		chain_id: The chain in the pdb you are trying to extract
+		start: The beginning residue that you want to start your extraction from
+		end:
[…]

### 2015-05-21 scripts/create_figures.R
Adding scripts but still needs more organization

@@ -1,0 +1,81 @@
+library(dplyr)
+library(ggplot2)
+
+correlation_data = read.csv("../data/all_protein_summary_stats.csv")
+
+p1 = ggplot(correlation_data, aes(x = mean_entropy, y = rsa_entropy_rho, color = data_id)) +
+	theme_classic() + 
+  	xlab("Mean Entropy") + 
+  	ylab("Correlation RSA-Entropy") + 
+  	theme(text = element_text(size = 18)) + 
+  	theme(axis.text.x = element_text(face="bold")) + 
+  	theme(axis.line = element_line(size = 1.25)) + 
+  	theme(axis.ticks = element_line(size = 1.25)) + 
+  	geom_point(size = 2) 
+  	
+  	ggsave("../figures/mean_entropy_rsa_cor.pdf", plot  = p1, height = 4, width = 6)
+
+p2 = ggplot(correlation_data, aes(x = var_entropy, y = rsa_entropy_rho, color = data_id)) +
+	theme_classic() + 
+  	xlab("Variance Entropy") + 
+  	ylab("Correlation RSA-Entropy") + 
+  	 theme(text = element_text(size = 18)) + 
+  	theme(axis.text.x = element_text(face="bold")) + 
+  	theme(axis.line = element_line(size = 1.25)) + 
+  	theme(axis.ticks = element_line(size = 1.25)) + 
+  	geom_point(size = 2) 
+  	
+  	ggsave("../figures/var_entropy_rsa_cor.pdf", plot = p2, height = 4, width = 6)
+  
+
+p3 = ggplot(correlation_data, aes(x = mean_entropy, y = cn_entropy_rho, color = data_id)) +
+	theme_classic() + 
+  	xlab("Mean Entropy") + 
+  	ylab("Correlation CN-Entropy") + 
+  	theme(text = element_text(size = 18)) + 
+  	theme(axis.text.x = element_text(face="bold")) + 
+  	theme(axis.line = element_line(size = 1.25)) + 
+  	theme(axis.ticks = element_line(size = 1.25)) + 
+  	geom_point(size = 2) 
+  	 
+  	ggsave("../figures/mean_entropy_cn_cor.pdf", plot  = p3, height = 4, width = 6)
+  	
+
+p4 = ggplot(correlation_data, aes(x = var_entropy, y = cn_entropy_rho, color = data_id)) +
+	theme_classic() + 
+  	xlab("Variance Entropy") + 
[…]

### 2015-05-21 scripts/get_data.R

@@ -1,6 +1,98 @@
+library(plyr)
 library(dplyr)
+library(tidyr)
 enzyme_data = read.table(file = "../data/jec_ddg_209_monomers.csv", sep = ",", header = T)
-enzyme_out_data = data.frame(enzyme_data$pdb, enzyme_data$nsites, enzyme_data$nseq, enzyme_data$site, enzyme_data$rsa.tien, enzyme_data$r4s.jc.ej, enzyme_data$r4s.jtt.ej)
-enzyme_out_data = rename(enzyme_out_data,  pdb_id = enzyme_data.pdb, num_sites = enzyme_data.nsites, num_seqs = enzyme_data.nseq, site =  enzyme_data.site, rsa = enzyme_data.rsa.tien, r4c_jc = enzyme_data.r4s.jc.ej, r4s.jtt = enzyme_data.r4s.jtt.ej)
-enzyme_out_data = mutate(enzyme_out_data, data_id = "enzyme")
-write.csv(enzyme_out_data, "../data/huang_data.csv")
+enzyme_out_data = data.frame(enzyme_data$pdb, enzyme_data$site, enzyme_data$rsa.tien) #, enzyme_data$r4s.jc.ej, enzyme_data$r4s.jtt.ej)
+
+names(enzyme_out_data) <- c("pdb_id", "site", "rsa")  #, "cn", "wcn")
+
+mono_data <- read.table("../data/profiles_213_monomers.csv", sep  = ",", header = T)
+enzyme_contacts <- mono_data %>% select(pdb, site, WCN, CN.13A.)
+names(enzyme_contacts) <- c("pdb_id", "site", "cn", "wcn")
+final_enzyme_data = inner_join(enzyme_contacts, enzyme_out_data)
+
+final_enzyme_data = mutate(final_enzyme_data, data_id = "enzyme")
+write.csv(final_enzyme_data, "../data/huang_et_al_data.csv")
+
+entropy_data = read.table("../data/pdb_entropies.csv", sep = ",", header = T)
+
+new_entropy_data = entropy_data %>% select(pdb, site, entropy_from_alignments)
+names(new_entropy_data) <- c("pdb_id", "site", "entropy")
+reduced_entropy_data = new_entropy_data %>% filter(pdb_id != "1BBS" | pdb_id != "1BS0" | pdb_id != "1DIN" | pdb_id != "1HPL")
+write.csv(reduced_entropy_data, "../data/reduced_entropy_data.csv")
+
[…]

### 2015-05-21 scripts/get_entropies.R

@@ -1,0 +1,66 @@
+library(dplyr)
+HP_entropies = read.table("../entropies/entropies/HP_entropies.txt", sep = "\t",  header = T) #1RD8_AB
+HP_map = read.table("../entropies/pdb_maps/1RD8_AB_HP_map.txt", sep = "\t", header = T)
+HP_data = cbind(HP_entropies, HP_map) 
+mod_HP_data = HP_data %>% filter(pdb_pos != "NA")
+out_HP_data = mod_HP_data %>% select(res_num, entropy)
+write.csv(out_HP_data, "../data/mapped_entropies/1RD8_AB_entropies.csv")
+
+WNPB_entropies = read.table("../entropies/entropies/WNPB_entropies.txt", sep = "\t",  header = T) #2FP7_B
+WNPB_map = read.table("../entropies/pdb_maps/2FP7_B_WNPB_map.txt", sep = "\t", header = T)
+WNPB_data = cbind(WNPB_entropies, WNPB_map) 
+mod_WNPB_data = WNPB_data %>% filter(pdb_pos != "NA")
+out_WNPB_data = mod_WNPB_data %>% select(res_num, entropy)
+write.csv(out_WNPB_data, "../data/mapped_entropies/2FP7_B_entropies.csv")
+
+CCHFN_entropies = read.table("../entropies/entropies/CCHFN_entropies.txt", sep = "\t",  header = T) #4AQF_B
+CCHFN_map = read.table("../entropies/pdb_maps/4AQF_B_CCHFN_map.txt", sep = "\t", header = T)
+CCHFN_data = cbind(CCHFN_entropies, CCHFN_map) 
+mod_CCHFN_data = CCHFN_data %>% filter(pdb_pos != "NA")
+out_CCHFN_data = mod_CCHFN_data %>% select(res_num, entropy)
+write.csv(out_CCHFN_data, "../data/mapped_entropies/4AQF_B_entropies.csv")
+
+DPH_entropies = read.table("../entropies/entropies/DPH_entropies.txt", sep = "\t",  header = T) #2JLY_A
+DPH_map = read.table("../entropies/pdb_maps/2JLY_A_DPH_map.txt", sep = "\t", header = T)
+DPH_data = cbind(DPH_entropies, DPH_map) 
+mod_DPH_data = DPH_data %>% filter(pdb_pos != "NA")
+out_DPH_data = mod_DPH_data %>% select(res_num, entropy)
[…]

### 2015-05-21 scripts/summarize_protein_data.R

@@ -1,0 +1,57 @@
+library(plyr)
+library(dplyr)
+enzyme_data  = read.csv("../data/all_huang_et_al_data.csv")
+virus_data = read.csv("../data/sharamoradi_et_al_data.csv")
+
+pdb_ids = levels(virus_data$pdb_id)
+virus_summary = data.frame(pdb_id = character(0), cn_entropy_rho =numeric(0),cn_entropy_p_value = numeric(0), wcn_entropy_rho =numeric(0), wcn_entropy_p_value = numeric(0), rsa_entropy_rho = numeric(0),rsa_entropy_p_value = numeric(0), mean_entropy = numeric(0), var_entropy = numeric(0), sd_entropy = numeric(0))
+
+for(i in pdb_ids){
+
+	data = filter(virus_data, pdb_id == i)
+	r_cn = as.numeric(cor.test(data$entropy, data$cn, method = "spearman")$estimate)
+	p_cn = cor.test(data$entropy, data$cn, method = "spearman")$p.value
+	
+	r_wcn = as.numeric(cor.test(data$entropy, data$wcn, method = "spearman")$estimate)
+	p_wcn = cor.test(data$entropy, data$wcn, method = "spearman")$p.value
+	
+	r_rsa = as.numeric(cor.test(data$entropy, data$rsa, method = "spearman")$estimate)
+	p_rsa = cor.test(data$entropy, data$rsa, method = "spearman")$p.value
+	
+	mean_ent = mean(data$entropy)
+	var_ent = var(data$entropy)
+	sd_ent = sd(data$entropy)
+	
+	new_data = data.frame(pdb_id = i, cn_entropy_rho = r_cn, cn_entropy_p_value = p_cn, wcn_entropy_rho = r_wcn, wcn_entropy_p_value = p_wcn, rsa_entropy_rho = r_rsa, rsa_entropy_p_value = p_rsa, mean_entropy = mean_ent, var_entropy = var_ent, sd_entropy = sd_ent)
+	virus_summary = rbind(virus_summary, new_data)
+}
+
+pdb_ids = levels(enzyme_data$pdb_id)
[…]

### 2015-05-21 manuscript/manuscript.tex
Change name of tex

--- 2015-05-16 manuscript/manuscript_cordiv.tex
@@
 
-\graphicspath{{../../analysis/figures/}}
+\graphicspath{{../figures/}}
 
@@
 \bibliographystyle{peerj} %"style
-\bibliography{ej_manuscript_bib} %expected file "my refs.bib"
+\bibliography{manuscript_bib} %expected file "my refs.bib"
 
+\cleardoublepage
 \section*{Figures}
 
-1. WCN-Rate4site vs Variance of RateSite (Also for RSA) and CN-Rate4site vs Variance of RateSite \\
-2. WCN-Entropy vs Mean Entropy  and CN-Entropy vs Mean Entropy \\
-3. RSA-Entropy vs Mean Entropy \\
-4. RSA vs Rate4Site \\
-5. Should I make the stress  and delta delta G correlations plots for the viral sequences that would complete the story?
+    \begin{figure}[H]
+            \centerline{\includegraphics[width=4in]{mean_entropy_cn_cor.pdf}}     
+            \caption{NEED A CAPTION!!!.}
+            \label{fig:seqent_structure_cors}
+    \end{figure}
 
-%    \begin{figure}[tbh]
-%            \begin{center}
-%            %\includegraphics[width=6.9in]{cordiv_similarities_r4sJC.pdf}
-%            \includegraphics[width=6.9in]{cordiv_similarities_seqent.pdf}
-%            \end{center}
-%            \caption{A comparison of the strength of Spearman correlation of sequence evolutionary rates (r4sJC) with {\it side chain} Weighted Contact Number vs.
-correlations of other structural properties with evolutionary rates.
-Detailed description of the structural properties is given in Section \label{sec:mam}.
-The red lines in each plot represent equality line.
-It is evident from all plots that for any given protein in dataset, the correlation strength of one structural property is a good proxy measure of the correlation strength of any other structural property with sequence variability measures.
-For brevity, correlations of structure-rate4site are not shown here but are available online, also in supplementary material.}
-%            \label{fig:seqent_structure_cors}
-%    \end{figure}
-%
-%
-%    \begin{figure}[tbh]
-%        \centering
-%        \includegraphics[width=3.6in]{/main_modulators_cormat_r4s_wcn.pdf} \\
-%        \caption{{\bf The Spearman correlation matrix of the strongest sequence--structure correlations and the prominent determinants of the strengths of the corresponding relations}.
-The variables on the diagonal elements of the matrix from top to bottom represent respectively, the four strongest sequence--structure relations: r4sJC--wcnSC, r4sJC--varea, seqent--wcnSC, seqent--varea, followed by protein properties that modulate the strength of these relations: the correlation between r4sJC--seqent ({\it r.r4sJC.seqent}), variance of r4sJC ({\it var.r4sJC}), variance of sequence entropy ({\it var.seqent}), variance of back-bone hydrogen bond energies ({\it var.hbe}), and the fraction of amino acids in helical \& $\beta-$sheet secondary structures in the proteins ({\it mn.helix} \& {\it mn.betas} respectively).}
-%        \label{fig:mainmods}
-%    \end{figure}
+    \begin{figure}[H]
+            \centerline{\includegraphics[width=4in]{var_entropy_cn_cor.pdf}}
+            \caption{NEED A CAPTION!!!.}
+            \label{fig:seqent_structure_cors}
+    \end{figure}
 
-%    \begin{figure*}[tbh]
-  %      \centering
- %       \includegraphics[width=3.4in]{/validation_r_r4s_wcn_var_seqent.pdf}
- %       \caption{{\bf Sequence--structure correlation strength versus sequence divergence.} The plot illustrates the relationship between the strength of a representative sequence--structure correlation ({\it seqent--scnSC}) and the sequence divergence as measured by the variance of protein sequence entropy.
-The black circles represent $209$ proteins used in this work.
-For comparison and validation, the red circles represent data for $9$ viral proteins taken from {\color{red}shahmoradi2014xx}}.
-%        \label{fig:validation}
-%    \end{figure*}
+    \begin{figure}[H]
+            \centerline{\includegraphics[width=4in]{mean_entropy_wcn_cor.pdf}}
+            \caption{NEED A CAPTION!!!.}
+            \label{fig:seqent_structure_cors}
+    \end{figure}
 
-%    \begin{figure}[tbh]
-%        \centering
-%        \includegraphics[width=7in]{/best_wcn/select_variables/boxplot_wcn_all_in_one.pdf} \\
-%        \caption{NEEDS A CAPTION.}
-%        \label{fig:best_wcn}
-%    \end{figure}
-%
-%    \begin{figure*}[tbh]
-%        \centering
-%        \begin{tabular}{cc}
-%        \includegraphics[width=3.4in]{/modulators_similarity_Sr4sJC_scors.pdf} & \includegraphics[width=3.4in]{/modulators_similarity_Sseqent_scors.pdf} \\
-%        \end{tabular}
-%        \caption{{\bf Spearman correlation matrices illustrating the similarity of the set of structure and sequence properties that contribute to the strength of sequence--structure relations}.
-{\bf Left:} The correlation matrix for the relation of evolutionary rates ({\it r4sJC}) with structural quantities on the diagonal of the matrix.
-{\bf Right:} The correlation matrix for the relation of sequence entropy ({\it seqent}) with structural quantities on the diagonal of the matrix.}
-%        \label{fig:simcormats}
-%    \end{figure*}
-%
-%    \begin{figure*}[tbh]
-%        \centering
-%        %\begin{tabular}{cc}
-%        \includegraphics[width=3.4in]{/validation_r_seqent_wcn_var_seqent.pdf}
-%        %\end{tabular}
-%        \caption{{\bf Sequence--structure correlation strength versus sequence divergence.} The plot illustrates the relationship between the strength of a representative sequence--structure correlation ({\it seqent--scnSC}) and the sequence divergence as measured by the variance of protein sequence entropy.
-The black circles represent $209$ proteins used in this work.
-For comparison and validation, the red circles represent data for $9$ viral proteins taken from {\color{red} shahmoradi2014xx}}.
-%        \label{fig:validation}
-%    \end{figure*}
+    \begin{figure}[H]
+            \centerline{\includegraphics[width=4in]{var_entropy_wcn_cor.pdf}}
+            \caption{NEED A CAPTION!!!.}
+            \label{fig:seqent_structure_cors}
+    \end{figure}
+
+    \begin{figure}[H]
+            \centerline{\includegraphics[width=4in]{mean_entropy_rsa_cor.pdf}}
+            \caption{NEED A CAPTION!!!.}
+            \label{fig:seqent_structure_cors}
+    \end{figure}
+
+    \begin{figure}[H]
+            \centerline{\includegraphics[width=4in]{var_entropy_rsa_cor.pdf}}
+            \caption{NEED A CAPTION!!!.}
+            \label{fig:seqent_structure_cors}
+    \end{figure}
+
+NEED THE RATE4SITE FIGURES
 

### 2015-05-23 scripts/get_data.R
Organizing repo

@@ -1,98 +1,135 @@
 library(plyr)
 library(dplyr)
 library(tidyr)
-enzyme_data = read.table(file = "../data/jec_ddg_209_monomers.csv", sep = ",", header = T)
-enzyme_out_data = data.frame(enzyme_data$pdb, enzyme_data$site, enzyme_data$rsa.tien) #, enzyme_data$r4s.jc.ej, enzyme_data$r4s.jtt.ej)
+enzyme_data = read.table(file = "../data/enzyme_dataset/jec_ddg_209_monomers.csv", sep = ",", header = T)
+enzyme_out_data = data.frame(enzyme_data$pdb, enzyme_data$site, enzyme_data$rsa.tien, enzyme_data$r4s.jtt.ej)
 
-names(enzyme_out_data) <- c("pdb_id", "site", "rsa")  #, "cn", "wcn")
+names(enzyme_out_data) <- c("pdb_id", "site", "rsa", "r4s_JTT")
 
-mono_data <- read.table("../data/profiles_213_monomers.csv", sep  = ",", header = T)
+mono_data <- read.table("../data/enzyme_dataset/profiles_213_monomers.csv", sep  = ",", header = T)
 enzyme_contacts <- mono_data %>% select(pdb, site, WCN, CN.13A.)
 names(enzyme_contacts) <- c("pdb_id", "site", "cn", "wcn")
 final_enzyme_data = inner_join(enzyme_contacts, enzyme_out_data)
 
 final_enzyme_data = mutate(final_enzyme_data, data_id = "enzyme")
-write.csv(final_enzyme_data, "../data/huang_et_al_data.csv")
+write.csv(final_enzyme_data, "../data/output_files/enzyme_data.csv")
 
-entropy_data = read.table("../data/pdb_entropies.csv", sep = ",", header = T)
+entropy_data = read.table("../data/enzyme_dataset/pdb_entropies.csv", sep = ",", header = T)
 
 new_entropy_data = entropy_data %>% select(pdb, site, entropy_from_alignments)
 names(new_entropy_data) <- c("pdb_id", "site", "entropy")
 reduced_entropy_data = new_entropy_data %>% filter(pdb_id != "1BBS" | pdb_id != "1BS0" | pdb_id != "1DIN" | pdb_id != "1HPL")
-write.csv(reduced_entropy_data, "../data/reduced_entropy_data.csv")
[…]

### 2015-05-24 evol_rates/scripts/extract_rate4site_rates.py
Adding files

@@ -1,0 +1,113 @@
+import protein_toolbox_helper as ph
+import numpy as np
+import re, os, sys
+
+'''
+Last Edited By: Eleisha Jackson on February 9, 2015
+This is a script that extracts the amino acid rates from the individual Rate4Site file for each protein and then places it into a file.
+'''
+
+def get_rates(map_file, rate_file):
+	'''
+	This is a function that extracts and maps the evolutionary rates to the pdb sequence
+	
+	Args:
+
+		map_file: The file that maps the pdb sequence to the alignment	
+		rate_file: The rate4site file
+	Returns:
+		sites: The sites in the pdb that have been mapped to the alignment	
+		aligned_rates: The rates mapped back to sites in the pdb	
+	'''
+			
+	align_sites = []
+	rates = []
+	amino_acids = []
+	ref_seq_array  = []
+
+	input = open(rate_file)
+	data = input.readlines()
+	data = data[11:]
+	data.pop()
+	data.pop()
+
+	for line in data: #Get the evolutionary rates from the rate4site file
+		site = line[2:8].strip()
+		amino = line[9:12].strip()
+		rate = line[11:19].strip()
+		align_sites.append(site)
+		amino_acids.append(amino)
+		rates.append(rate)	
+
+	sites = extract_sites(map_file)
+	aligned_rates = []
+	counter = 0
+	ref_seq_array = np.genfromtxt(map_file, delimiter = "\t", dtype = "string", skip_header = 1, usecols = (2)) #Read in the data
+	#print ref_seq_array
+	#Align the calculated rates to the proper amnio acid in the pdb using the pdb_to_alignment_map created earlier (in match_frequencies.py)
+	
+	for j in xrange(0, len(ref_seq_array)):		
+		if(ref_seq_array[j] == '-'):
+			#print str(j+1), ref_seq_array[j], sites[j], "NA"
+			aligned_rates.append("NA")
+		else:
+			#print str(j+1), ref_seq_array[j], sites[j]
+			aligned_rates.append(rates[counter])
+			counter = counter + 1
+	return sites, aligned_rates
+
+def extract_sites(map_file):
+	'''
+	This is a function extracts the site information from the mapped file
+	
+	Args:
+		pdb: The pdb name
+		map_file: The file that maps the pdb sequence to the alignment
+		rate_file: the rate4site file
+	Returns:
+		sites: The sites in the pdb that have been mapped to the alignment		
+	'''
+
[…]

### 2015-05-25 manuscript/manuscript.tex
More figures

@@
 
-\title{Sequence divergence as the main determinant of sequence-structure relationships}
+\title{Sequence variability as the main determinant of sequence-structure relationships}
 \author{Amir Shahmoradi$^{1*}$, Eleisha L. Jackson$^2$, Claus O. Wilke$^2$}
@@
     \begin{figure}[H]
-            \centerline{\includegraphics[width=4in]{mean_entropy_cn_cor.pdf}}     
+            \centerline{\includegraphics[width=7.5in]{entropy_cn_cor.pdf}}     
             \caption{NEED A CAPTION!!!.}
@@
     \begin{figure}[H]
-            \centerline{\includegraphics[width=4in]{var_entropy_cn_cor.pdf}}
+            \centerline{\includegraphics[width=7.5in]{entropy_wcn_cor.pdf}}     
             \caption{NEED A CAPTION!!!.}
@@
     \begin{figure}[H]
-            \centerline{\includegraphics[width=4in]{mean_entropy_wcn_cor.pdf}}
+            \centerline{\includegraphics[width=7.5in]{entropy_rsa_cor.pdf}}     
             \caption{NEED A CAPTION!!!.}
@@
         \begin{figure}[H]
-            \centerline{\includegraphics[width=4in]{var_entropy_wcn_cor.pdf}}
+            \centerline{\includegraphics[width=7.5in]{rate_cor.pdf}}     
             \caption{NEED A CAPTION!!!.}
@@
             \begin{figure}[H]
-            \centerline{\includegraphics[width=4in]{mean_entropy_rsa_cor.pdf}}
+            \centerline{\includegraphics[width=5.0in]{var_rate_rsa_cor.pdf}}     
             \caption{NEED A CAPTION!!!.}
@@
         \begin{figure}[H]
-            \centerline{\includegraphics[width=4in]{var_entropy_rsa_cor.pdf}}
+            \centerline{\includegraphics[width=5.0in]{var_entropy_rate_cor.pdf}}     
             \caption{NEED A CAPTION!!!.}
@@
 
-NEED THE RATE4SITE FIGURES
-
 \end{document}

### 2015-05-25 scripts/create_figures.R

@@ -1,9 +1,12 @@
 library(dplyr)
 library(ggplot2)
+library(cowplot)
 
-correlation_data = read.csv("../data/all_protein_summary_stats.csv")
+correlation_data = read.csv("../data_summaries/output_files/all_protein_summary_stats.csv")
+correlation_data = rename(correlation_data, dataset = data_id)
+enzyme_data = filter(correlation_data, dataset == "enzyme")
 
-p1 = ggplot(correlation_data, aes(x = mean_entropy, y = rsa_entropy_rho, color = data_id)) +
+p1 = ggplot(correlation_data, aes(x = mean_entropy, y = rsa_entropy_rho, color = dataset)) +
 	theme_classic() + 
   	xlab("Mean Entropy") + 
   	ylab("Correlation RSA-Entropy") + 
@@ -11,11 +14,13 @@ p1 = ggplot(correlation_data, aes(x = mean_entropy, y = rsa_entropy_rho, color =
   	theme(axis.text.x = element_text(face="bold")) + 
   	theme(axis.line = element_line(size = 1.25)) + 
   	theme(axis.ticks = element_line(size = 1.25)) + 
-  	geom_point(size = 2) 
-  	
-  	ggsave("../figures/mean_entropy_rsa_cor.pdf", plot  = p1, height = 4, width = 6)
+  	geom_point(size = 2) + 
+  	stat_smooth(method = "lm", se = F, data = enzyme_data) + 
+  	scale_color_manual(values = c("black", "red"))
+  	  	
+  	ggsave("../figures/mean_entropy_rsa_cor.pdf", plot = p1, height = 4.5, width = 6)
 
-p2 = ggplot(correlation_data, aes(x = var_entropy, y = rsa_entropy_rho, color = data_id)) +
+p2 = ggplot(correlation_data, aes(x = var_entropy, y = rsa_entropy_rho, color = dataset)) +
 	theme_classic() + 
   	xlab("Variance Entropy") + 
   	ylab("Correlation RSA-Entropy") + 
@@ -23,12 +28,18 @@ p2 = ggplot(correlation_data, aes(x = var_entropy, y = rsa_entropy_rho, color = 
   	theme(axis.text.x = element_text(face="bold")) + 
   	theme(axis.line = element_line(size = 1.25)) + 
   	theme(axis.ticks = element_line(size = 1.25)) + 
[…]

### 2015-05-25 scripts/get_data.R

@@ -1,135 +1,119 @@
 library(plyr)
 library(dplyr)
 library(tidyr)
-enzyme_data = read.table(file = "../data/enzyme_dataset/jec_ddg_209_monomers.csv", sep = ",", header = T)
+
+enzyme_data = read.table(file = "../data_summaries/enzyme_dataset/jec_ddg_209_monomers.csv", sep = ",", header = T)
 enzyme_out_data = data.frame(enzyme_data$pdb, enzyme_data$site, enzyme_data$rsa.tien, enzyme_data$r4s.jtt.ej)
 
 names(enzyme_out_data) <- c("pdb_id", "site", "rsa", "r4s_JTT")
 
-mono_data <- read.table("../data/enzyme_dataset/profiles_213_monomers.csv", sep  = ",", header = T)
+mono_data <- read.table("../data_summaries/enzyme_dataset/profiles_213_monomers.csv", sep  = ",", header = T)
 enzyme_contacts <- mono_data %>% select(pdb, site, WCN, CN.13A.)
 names(enzyme_contacts) <- c("pdb_id", "site", "cn", "wcn")
 final_enzyme_data = inner_join(enzyme_contacts, enzyme_out_data)
 
 final_enzyme_data = mutate(final_enzyme_data, data_id = "enzyme")
-write.csv(final_enzyme_data, "../data/output_files/enzyme_data.csv")
+#write.csv(final_enzyme_data, "../data_summaries/output_files/enzyme_data.csv")
 
-entropy_data = read.table("../data/enzyme_dataset/pdb_entropies.csv", sep = ",", header = T)
+entropy_data = read.table("../data_summaries/enzyme_dataset/pdb_entropies.csv", sep = ",", header = T)
 
 new_entropy_data = entropy_data %>% select(pdb, site, entropy_from_alignments)
 names(new_entropy_data) <- c("pdb_id", "site", "entropy")
 reduced_entropy_data = new_entropy_data %>% filter(pdb_id != "1BBS" | pdb_id != "1BS0" | pdb_id != "1DIN" | pdb_id != "1HPL")
-write.csv(reduced_entropy_data, "../data/output_files/enzyme_reduced_entropy_data.csv")
+#write.csv(reduced_entropy_data, "../data_summaries/output_files/enzyme_reduced_entropy_data.csv")
 
[…]

### 2015-05-25 scripts/get_entropies.R

@@ -1,66 +1,63 @@
 library(dplyr)
 HP_entropies = read.table("../entropies/entropies/HP_entropies.txt", sep = "\t",  header = T) #1RD8_AB
-HP_map = read.table("../entropies/pdb_maps/1RD8_AB_HP_map.txt", sep = "\t", header = T)
+HP_map = read.table("../pdb_maps/1RD8_AB_HP_map.txt", sep = "\t", header = T)
 HP_data = cbind(HP_entropies, HP_map) 
 mod_HP_data = HP_data %>% filter(pdb_pos != "NA")
 out_HP_data = mod_HP_data %>% select(res_num, entropy)
-write.csv(out_HP_data, "../data/mapped_entropies/1RD8_AB_entropies.csv")
+write.csv(out_HP_data, "../data_summaries/virus_dataset/mapped_entropies/1RD8_AB_entropies.csv")
 
 WNPB_entropies = read.table("../entropies/entropies/WNPB_entropies.txt", sep = "\t",  header = T) #2FP7_B
-WNPB_map = read.table("../entropies/pdb_maps/2FP7_B_WNPB_map.txt", sep = "\t", header = T)
+WNPB_map = read.table("../pdb_maps/2FP7_B_WNPB_map.txt", sep = "\t", header = T)
 WNPB_data = cbind(WNPB_entropies, WNPB_map) 
 mod_WNPB_data = WNPB_data %>% filter(pdb_pos != "NA")
 out_WNPB_data = mod_WNPB_data %>% select(res_num, entropy)
-write.csv(out_WNPB_data, "../data/mapped_entropies/2FP7_B_entropies.csv")
+write.csv(out_WNPB_data, "../data_summaries/virus_dataset/mapped_entropies/2FP7_B_entropies.csv")
 
 CCHFN_entropies = read.table("../entropies/entropies/CCHFN_entropies.txt", sep = "\t",  header = T) #4AQF_B
-CCHFN_map = read.table("../entropies/pdb_maps/4AQF_B_CCHFN_map.txt", sep = "\t", header = T)
+CCHFN_map = read.table("../pdb_maps/4AQF_B_CCHFN_map.txt", sep = "\t", header = T)
 CCHFN_data = cbind(CCHFN_entropies, CCHFN_map) 
 mod_CCHFN_data = CCHFN_data %>% filter(pdb_pos != "NA")
 out_CCHFN_data = mod_CCHFN_data %>% select(res_num, entropy)
[…]

### 2015-05-25 scripts/get_rates.R

@@ -1,0 +1,66 @@
+library(dplyr)
+HP_rates = read.table("../evol_rates/rate4site/rates/unnormalized/1RD8_AB_evol_rates.txt", sep = ",",  header = T) #1RD8_AB
+HP_map = read.table("../pdb_maps/1RD8_AB_HP_map.txt", sep = "\t", header = T)
+HP_data = cbind(HP_rates, HP_map) 
+mod_HP_data = HP_data %>% filter(pdb_pos != "NA")
+out_HP_data = mod_HP_data %>% select(site, rate)
+write.csv(out_HP_data, "../data_summaries/virus_dataset/mapped_rates/1RD8_AB_rates.csv")
+
+WNPB_rates = read.table("../evol_rates/rate4site/rates/unnormalized/2FP7_B_evol_rates.txt", sep = ",",  header = T) #2FP7_B
+WNPB_map = read.table("../pdb_maps/2FP7_B_WNPB_map.txt", sep = "\t", header = T)
+WNPB_data = cbind(WNPB_rates, WNPB_map) 
+mod_WNPB_data = WNPB_data %>% filter(pdb_pos != "NA")
+out_WNPB_data = mod_WNPB_data %>% select(site, rate)
+write.csv(out_WNPB_data, "../data_summaries/virus_dataset/mapped_rates/2FP7_B_rates.csv")
+
+CCHFN_rates = read.table("../evol_rates/rate4site/rates/unnormalized/4AQF_B_evol_rates.txt", sep = ",",  header = T) #4AQF_B
+CCHFN_map = read.table("../pdb_maps/4AQF_B_CCHFN_map.txt", sep = "\t", header = T)
+CCHFN_data = cbind(CCHFN_rates, CCHFN_map) 
+mod_CCHFN_data = CCHFN_data %>% filter(pdb_pos != "NA")
+out_CCHFN_data = mod_CCHFN_data %>% select(site, rate)
+write.csv(out_CCHFN_data, "../data_summaries/virus_dataset/mapped_rates/4AQF_B_rates.csv")
+
+DPH_rates = read.table("../evol_rates/rate4site/rates/unnormalized/2JLY_A_evol_rates.txt", sep = ",",  header = T) #2JLY_A
+DPH_map = read.table("../pdb_maps/2JLY_A_DPH_map.txt", sep = "\t", header = T)
+DPH_data = cbind(DPH_rates, DPH_map) 
[…]

### 2015-05-25 scripts/summarize_protein_data.R

@@ -1,10 +1,10 @@
 library(plyr)
 library(dplyr)
-enzyme_data  = read.csv("../data/all_huang_et_al_data.csv")
-virus_data = read.csv("../data/sharamoradi_et_al_data.csv")
+enzyme_data  = read.csv("../data_summaries/output_files/all_enzyme_data.csv")
+virus_data = read.csv("../data_summaries/output_files/all_virus_data.csv")
 
 pdb_ids = levels(virus_data$pdb_id)
-virus_summary = data.frame(pdb_id = character(0), cn_entropy_rho =numeric(0),cn_entropy_p_value = numeric(0), wcn_entropy_rho =numeric(0), wcn_entropy_p_value = numeric(0), rsa_entropy_rho = numeric(0),rsa_entropy_p_value = numeric(0), mean_entropy = numeric(0), var_entropy = numeric(0), sd_entropy = numeric(0))
+virus_summary = data.frame(pdb_id = character(0), cn_entropy_rho = numeric(0),cn_entropy_p_value = numeric(0), wcn_entropy_rho =numeric(0), wcn_entropy_p_value = numeric(0), rsa_entropy_rho = numeric(0),rsa_entropy_p_value = numeric(0), rate_wcn_rho = numeric(0), rate_wcn_p_value = numeric(0), rate_cn_rho = numeric(0), rate_cn_p_value = numeric(0), rate_rsa_rho = numeric(0), rate_rsa_p_value = numeric(0), mean_entropy = numeric(0), var_entropy = numeric(0), sd_entropy = numeric(0), var_rate = numeric(0), sd_rate = numeric(0))
 
 for(i in pdb_ids){
 
@@ -17,17 +17,32 @@ for(i in pdb_ids){
 	
 	r_rsa = as.numeric(cor.test(data$entropy, data$rsa, method = "spearman")$estimate)
 	p_rsa = cor.test(data$entropy, data$rsa, method = "spearman")$p.value
+
+	r2_cn = as.numeric(cor.test(data$r4s_JTT, data$cn, method = "spearman")$estimate)
+	p2_cn = cor.test(data$r4s_JTT, data$cn, method = "spearman")$p.value
+	
+	r2_wcn = as.numeric(cor.test(data$r4s_JTT, data$wcn, method = "spearman")$estimate)
[…]

### 2015-05-25 manuscript/manuscript.tex
Manuscript changes

@@
 However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.
-Here we present the results from a comprehensive search for potential biophysical and structural determinants of protein evolution by studying more than $200$ structural and evolutionary properties in a dataset of $209$ monomeric enzymes.
-We discuss the main protein characteristics responsible for the general patterns of protein evolution, and identify sequence divergence as the main determinant of the strengths of virtually all structural-evolution relationships, explaining $10-30 \%$ of observed variation in sequence-structure relations.
-In addition to sequence divergence, we identify several protein structural properties that are moderately but significantly coupled with the strength of sequence-structure relations.
-In particular, proteins with more homogeneous back-bone hydrogen bond energies, large fractions of helical secondary structures and low fraction of beta sheets tend to have the strongest correlations between structural properties and site variability.
-
 \end{abstract}
@@
 \\
-COPY EDIT THIS!!! \\
 
@@
     The results presented in this work are based on a two datasets.
-The first is a dataset of $209$ monomeric enzymes from \citep{Echaveetal2015} originally from Huang et al.
-(2014).The original dataset was comprised of $213$ but we removed four of the proteins (1BBS, 1BS0, 1DIN, 2HPL) that had did not have data at insertion sites.
-Briefly, these proteins are all enzyme monomers   randomly picked from the Catalytic Site Atlas $2.2.11$ \citep{Porteretal2004} with protein sizes in the sample ranging from $95$ to $1287$.
+The first is a dataset of $209$ monomeric enzymes from \cite{Huangetal2014}, originally from \cite{Yehetal2014}.The original dataset was comprised of $213$ but we removed four of the proteins (1BBS, 1BS0, 1DIN, 2HPL) that had did not have data at insertion sites.
+Briefly, these proteins are all enzyme monomers   randomly picked from the Catalytic Site Atlas $2.2.11$ \citep{Porteretal2004} with protein sizes in the sample ranging from $95$ to $1287$ residues in length.
 For each structure we had a corresponding alignment of up to 300 homologous sequences.
-The second dataset was from taken from Sharamoradi et al.
-and is comprised of nine?
-viral proteins.
-The viral proteins range from 122 - 557 in length and each structure is accompanied by a sequence alignment of up to 2362 homologous sequences.
+The second dataset was from taken from \cite{Shahmoradietal2014} and is comprised of nine viral proteins.
+The viral proteins range from 122 - 557 residues in length and each structure is accompanied by a sequence alignment of up to 2362 homologous sequences.
 Sequence alignments for both datasets constructed using amino-acid sequences with MAFFT \citep{Katohetal2002, Katohetal2005}, specifying the auto flag to select the optimal algorithm for the given data set.The alignments were then used to calculate the site-specific evolutionary rates for each individual protein in both datasets.
@@
 
-All data and analysis scripts required to reproduce the work are publicly available to view and download at \url{https://github.com/shahmoradi/cordiv}.
+All data and analysis scripts required to reproduce the work are publicly available to view and download at \url{https://github.com/wilkelab/rate_variability_variation}.
 

### 2015-05-26 manuscript/manuscript.tex
Manuscript changes

@@
 \title{Sequence variability as the main determinant of sequence-structure relationships}
-\author{Amir Shahmoradi$^{1*}$, Eleisha L. Jackson$^2$, Claus O. Wilke$^2$}
+\author{Amir Shahmoradi$^1$, Eleisha L. Jackson$^2$, Claus O. Wilke$^{2*}$}
 \begin{document}
@@
 $^*$Corresponding author\\
-$\phantom{^*}$Email: amir@physics.utexas.edu\\
+%$\phantom{^*}$Email: amir@physics.utexas.edu\\
+$\phantom{^*}$Email: wilke@austin.utexas.edu\\
 %$\phantom{^*}$Phone:{ \color{red} Need a phone} \\
@@
 \begin{abstract}
-NEED TO EDIT THE ABSTRACT\\
 Recent work has shown that structural properties are capable of predicting site-specific sequence variability for a given protein.
@@
 \\
-\indent Later, Yeh et al performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by CN and WCN \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
-A year later, Sharamoradi et al also performed a site-wise analysis on a series of viral proteins.
+\indent Later, \cite{Yehetal} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by CN and WCN \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
+A year later, \cite{Sharamoradietal} also performed a site-wise analysis on a series of viral proteins.
 In this analysis they found that RSA had the strongest correlation with site variability.
@@
 First the Maximum Likelihood phylogenetic trees were inferred with RAxML, using the LG substitution matrix and the CAT model of rate heterogeneity \citep{Stamatakis2006, Stamatakis2014}.
-For each structure, we then used the respective sequence alignment and phylogenetic tree to infer site-specific substitution rates with Rate4Site, using the empirical Bayesian method and the amino-acid Jukes-Cantor mutational model (aaJC) \citep{Mayroseetal2004}.
+For each structure, we then used the respective sequence alignment and phylogenetic tree to infer site-specific substitution rates with Rate4Site, using the empirical Bayesian method and the JJT model of sequence evolution \citep{Mayroseetal2004}.
 
     \subsection*{Calculation of Structural Properties}
+In our analysis we used to types of measures of local packing density used in previous studies: Contact Number and Weighted Contact Number).
+For the purposes of our comparison between the two datasets of interest we used the CN and WCN for the enzyme proteins from () and the CN and WCN values for the viral proteins from.
+In both of these works and in Yeh et al, WCN and CN are both defined the same.
+Contact Number is defined as the number of $C_{\alpha}$ within a pre-redefined radius, $r_0$.
+In this case, $r_0$ = 13 as in the previous work.
+Weighed Contact Number for a residue, $i$,  is defined as in (Liao et al 2005, Huang et al 2014), as:
 
-    HOW IS CN and WCN Defined and Calculated??? (Cite Shih et al for WCN) Possbily Lin Paper \\
+	\begin{equation} \label{wcn_eqn}
+		\text{WCN}_i = \sum_{i \neq j}^{N} \frac{1}{r_{ij}^2 } 
+	\end{equation}
+
+	where $r_j$ is the length between the $C_{\alpha}$ of residue $i$ and residue $j$ in a protein of length $N$.
+
 \indent We used DSSP \citep{Kabschetal2005}  for the calculation of Accessible Surface Area (ASA) for each site.
 We normalized the ASA for each site by the theoretical maximum solvent accessibility values of \cite{Tienetal2013}  to obtain the Relative Solvent Accessibility (RSA) for all individual sites in all proteins.
-In addition to ASA values, we also extract from DSSP output, information about the secondary structure of proteins such as the total number of residues participating in different types of helices, parallel or anti-parallel beta sheets, or loops and turns.
-To complete the list of pdb-level structural properties, we also calculate the Spearman correlations between all residue-level structure and sequence properties and include them in the analysis to probe their potential effects on the strength of structure-sequence relations.
 \\
@@
 
-    \subsection*{}
-{\color{red}SOME RESULTS.....} \\
+We calculate the strength between the structural properties and site variability by calculating the Spearman correlation between the structural property (solvent exposure and local packing density).
+On average the correlations between CN-Entropy are stronger in magnitude in the enzyme proteins as compared to viral proteins.
+The average rho between each structural property and each measure of site variability (entropy and evolutionary rates) can be seen in Table 1 and Table 2.
+The viral proteins experience much lower site variability.
+Figure 1 shows the correlations between CN and entropy for the virus and enzyme proteins.
+Even though on average the correlations in the enzyme proteins are larger, as reported, by \cite{Yehetal2014}, the virus proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations.
+The average site variability of a given protein does not seem to be a significant determinate of the strength of structural correlations (Figure 1A, 2A, 3A).
+However, when examining the variance of entropy there is a clear trend within the enzyme proteins.
+Proteins with a highr variance in city variability across the protein typically have lower correlations (Figure 1B, 2B, 3B) have lower correlations (Figure 1B, 2B, 3B).
+If you extrapoliate the trend from the enzyme proteins, the viral proteins follow a simllar trend.
+The correlation between the rho between RS and Entropy and the variance of entropy is positive.
+The correlation is negative between local packing density (both WCN and CN) and the variance of entropy.
+Unlike with entry, there is no relationship between variance of evolutionary rates at sites and any of the measured structural properties.
+There is a wide spread int he variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
 

### 2015-05-27 manuscript/manuscript.tex
Manuscript edits

--- 2015-05-25 manuscript/manuscript.tex
@@
 
-\title{Sequence variability as the main determinant of sequence-structure relationships}
-\author{Amir Shahmoradi$^{1*}$, Eleisha L. Jackson$^2$, Claus O. Wilke$^2$}
+\title{Variation in site entropy explains differences in structure-sequence relationships of proteins}
+\author{Amir Shahmoradi$^1$, Eleisha L. Jackson$^2$, Claus O. Wilke$^{2*}$}
 \begin{document}
@@
 \begin{abstract}
-NEED TO EDIT THE ABSTRACT\\
-Recent work has shown that structural properties are capable of predicting site-specific sequence variability for a given protein.
+{\color{red}Recent work has shown that structural properties are capable of predicting site-specific sequence variability for a given protein.
 However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.
+}
 \end{abstract}
@@
 Recently several site-specific structural properties that can explain patterns of sequence variability in proteins have been identified.
-One of the earliest examples is Relative Site Accessibility.
-\cite{Fransozaetal2009} identified RSA as the strongest predictor of evolutionary rate.
-They found that residues that are buried in the core of proteins tend to be more conserved than exposed residues close to the surface of the protein.
-In their analysis these concerned both RSA and various definitions of residue packing density to predict evolutionary rate.
+One of the earliest examples is Relative Site Accessibility (RSA).
+\cite{Fransozaetal2009} identified RSA as the strong predictor of evolutionary rate and found that residues that are buried in the core of proteins tend to be more conserved than exposed residues close to the surface of the protein.
+In their analysis they considered the ability of both RSA and various definitions of residue packing density to predict evolutionary rate.
 They found that RSA and evolutionary rate shared a significant linear relationship.
 Afterwards, several other works \citep{Ramseyetal2011, Scherreretal2012} also found that RSA as a very significant predictor of evolutionary rate and found this linear relationship as well.
-However in these papers all have the same flaw.
+However, these papers all have the same flaw.
 During the course of their analysis they binned the protein sites and average over all sites within a bin when looking at the trend of RSA.
-This could produce artifacts when looking at the trend.
+This process may have produce artifacts that account for this strong linear trend.
 \\
-\indent Later, Yeh et al performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by CN and WCN \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
-A year later, Sharamoradi et al also performed a site-wise analysis on a series of viral proteins.
-In this analysis they found that RSA had the strongest correlation with site variability.
+\indent Recently, \cite{Yehetal2014} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by CN and WCN \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
+A year later, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.
+In this analysis they found that RSA had the strongest correlation with site variability as opposed to local packing density.
 Additionally the effect seen between CN and WCN was of a much smaller magnitude as compared to \citep{Yehetal2014}.
+Therefore the relationship between local pacing density, RSA and site-specific measures of sequence variability is not well understood.
 Here we attempt to reconcile the work done in this area.
-We find that site variability is the primary determinant of structure determinant of the strength of structure-sequence relationships and some differences in previous work can be explained in terms of differing levels of site variability.
+We find that site variability is the primary determinant of the strength of structure-sequence relationships and some differences in previous work can be explained in terms of differing levels of site variability.
 \\
@@
     \subsection*{Structures, sequences, and measures of sequence properties } 
-    The results presented in this work are based on a two datasets.
-The first is a dataset of $209$ monomeric enzymes from \cite{Huangetal2014}, originally from \cite{Yehetal2014}.The original dataset was comprised of $213$ but we removed four of the proteins (1BBS, 1BS0, 1DIN, 2HPL) that had did not have data at insertion sites.
+    The results presented in this work are based on two datasets.
+The first is a dataset of $209$ monomeric enzymes taken from \cite{Huangetal2014}, originally from \cite{Yehetal2014}.The original dataset was comprised of $213$ proteins but we removed four of the proteins (PDB IDS: 1BBS, 1BS0, 1DIN, 2HPL) that had did not have data at insertion sites.
 Briefly, these proteins are all enzyme monomers   randomly picked from the Catalytic Site Atlas $2.2.11$ \citep{Porteretal2004} with protein sizes in the sample ranging from $95$ to $1287$ residues in length.
 For each structure we had a corresponding alignment of up to 300 homologous sequences.
-The second dataset was from taken from \cite{Shahmoradietal2014} and is comprised of nine viral proteins.
+The second dataset was taken from \cite{Shahmoradietal2014} and is comprised of nine viral proteins.
 The viral proteins range from 122 - 557 residues in length and each structure is accompanied by a sequence alignment of up to 2362 homologous sequences.
-Sequence alignments for both datasets constructed using amino-acid sequences with MAFFT \citep{Katohetal2002, Katohetal2005}, specifying the auto flag to select the optimal algorithm for the given data set.The alignments were then used to calculate the site-specific evolutionary rates for each individual protein in both datasets.
-To do so, we relied on two independent methods of measuring sequence variability measure.
-First, we calculated the Shannon entropy ($H_i$) -- the sequence entropy, hereafter abbreviated as {\it seqent} -- at each alignment column $i$:
+Sequence alignments for both datasets were constructed by aligning the amino-acid sequences using the alignment software MAFFT \citep{Katohetal2002, Katohetal2005}, specifying the auto flag to select the optimal algorithm for the given data set.
+The alignments were then used to calculate the site-specific measures of sequence variability for each individual protein in both datasets.
+To do so, we relied on two independent methods of measuring sequence variability.
+First, we calculated the Shannon entropy ($H_i$) -- the sequence entropy at each alignment column $i$:
     \begin{equation}
@@
 The sequence entropy is a measure of variability at each site.
-We also calculated a measure of site-specific evolutionary rate -- hereafter abbreviated as {\it r4s} -- for each protein using software Rate4site.
+We also calculated a measure of site-specific evolutionary rate for each protein using software Rate4site.
 First the Maximum Likelihood phylogenetic trees were inferred with RAxML, using the LG substitution matrix and the CAT model of rate heterogeneity \citep{Stamatakis2006, Stamatakis2014}.
-For each structure, we then used the respective sequence alignment and phylogenetic tree to infer site-specific substitution rates with Rate4Site, using the empirical Bayesian method and the amino-acid Jukes-Cantor mutational model (aaJC) \citep{Mayroseetal2004}.
+For each structure, we then used the respective sequence alignment and phylogenetic tree to infer site-specific substitution rates with Rate4Site using the empirical Bayesian method and the JTT model of sequence evolution \citep{Mayroseetal2004}.
 
     \subsection*{Calculation of Structural Properties}
+In our analysis we used two types of measures of local packing density used in previous studies: Contact Numbe r(CN) and Weighted Contact Number(WCN).
+For the purposes of our comparison between the two datasets of interest we used the CN and WCN for the enzyme proteins calculated in \cite{Huangetal2014} and the CN and WCN values for the viral proteins from \cite{Shahmoradietal2014}.
+In both of these works and in \cite{Yehetal2014}, WCN and CN are both defined the same.
+Contact Number is defined as the number of $C_{\alpha}$ within a pre-redefined radius, $r_0$.
+In this case, $r_0$ = 13 as in the previous papers.
+Weighed Contact Number for a residue, $i$, is defined as in (Liao et al 2005, Huang et al 2014), as:
 
-    HOW IS CN and WCN Defined and Calculated??? (Cite Shih et al for WCN) Possbily Lin Paper \\
-\indent We used DSSP \citep{Kabschetal2005}  for the calculation of Accessible Surface Area (ASA) for each site.
-We normalized the ASA for each site by the theoretical maximum solvent accessibility values of \cite{Tienetal2013}  to obtain the Relative Solvent Accessibility (RSA) for all individual sites in all proteins.
-In addition to ASA values, we also extract from DSSP output, information about the secondary structure of proteins such as the total number of residues participating in different types of helices, parallel or anti-parallel beta sheets, or loops and turns.
-To complete the list of pdb-level structural properties, we also calculate the Spearman correlations between all residue-level structure and sequence properties and include them in the analysis to probe their potential effects on the strength of structure-sequence relations.
+	\begin{equation} \label{wcn_eqn}
+		\text{WCN}_i = \sum_{i \neq j}^{N} \frac{1}{r_{ij}^2 } 
+	\end{equation}
+
+	where $r_j$ is the length between the $C_{\alpha}$ of residue $i$ and residue $j$ in a protein of length $N$ \citep{Yehetal2014}.
+
+\indent We used DSSP \citep{Kabschetal2005} to calculate the Accessible Surface Area (ASA) for each site.
+We then normalized the ASA for each site by the theoretical maximum solvent accessibility values of \cite{Tienetal2013} to obtain the Relative Solvent Accessibility (RSA) for all individual sites in all proteins.
 \\
@@
 
-    \subsection*{}
-{\color{red}SOME RESULTS.....} \\
+We calculate the strength between the structural properties and site variability by calculating the Spearman correlation, $\rho$,  between the structural property (solvent exposure and local packing density) and site variability, as measured by either evolutionary rate or site entropy.
+On average the correlations between CN and entropy are larger in absolute magnitude in the enzyme proteins as compared to viral proteins.
+The average $\rho$ between each structural property and each measure of site variability (entropy and evolutionary rates) can be seen in {\color{red}Table 1 and Table 2}.
+The viral proteins experience much lower site variability.
+Figure 1 shows the correlations between CN and entropy for the virus and enzyme proteins.
+Even though on average the correlations in the enzyme proteins are larger as reported, by \cite{Yehetal2014}, the virus proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations.
+
+\indent  The average site variability of a given protein does not seem to be a significant determinant of the strength of structural correlations (Figure 1A, 2A, 3A).
+However, when examining the variance of entropy there is a clear trend within the enzyme proteins.
+Proteins with a higher variance in site variability across the protein typically have have lower correlations (Figure 1B, 2B, 3B).
+If you extrapolate the trend from the enzyme proteins, the viral proteins follow a similar trend.
+The correlation between the $\rho$ between RSA and entropy and the variance of entropy is positive.
+The correlation is negative between local packing density (both WCN and CN) and the variance of entropy is negative.
+Unlike with entropy, there is no relationship between variance of evolutionary rates at sites and any of the measured structural properties.
+There also is a wide spread int he variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
 

### 2015-05-27 scripts/create_figures.R

@@ -10,9 +10,8 @@ p1 = ggplot(correlation_data, aes(x = mean_entropy, y = rsa_entropy_rho, color =
 	theme_classic() + 
   	xlab("Mean Entropy") + 
   	ylab("Correlation RSA-Entropy") + 
-  	theme(text = element_text(size = 18)) + 
-  	theme(axis.text.x = element_text(face="bold")) + 
-  	theme(axis.line = element_line(size = 1.25)) + 
+  	theme(text = element_text(size = 18)) +  #Get rid of bold
+  	theme(axis.line = element_line(size = 1.25)) + #Make this thinner
   	theme(axis.ticks = element_line(size = 1.25)) + 
   	geom_point(size = 2) + 
   	stat_smooth(method = "lm", se = F, data = enzyme_data) + 

### 2015-05-29 manuscript/manuscript.tex
Changed author list
Adding some comments on future edits

@@
 \title{Variation in site entropy explains differences in structure-sequence relationships of proteins}
-\author{Amir Shahmoradi$^1$, Eleisha L. Jackson$^2$, Claus O. Wilke$^{2*}$}
+\author{Eleisha L. Jackson$^{2^*}$, Amir Shahmoradi$^{1*}$, Claus O. Wilke$^{2^{**}}$}
 \begin{document}
@@
 
+\noindent $^{*}$These authors contributed equally. \\
+
 \bigskip
 \noindent
-$^*$Corresponding author\\
+$^{**}$Corresponding author\\
 %$\phantom{^*}$Email: amir@physics.utexas.edu\\
-$\phantom{^*}$Email: wilke@austin.utexas.edu\\
+$\phantom{^{** }}$Email: wilke@austin.utexas.edu\\
 %$\phantom{^*}$Phone:{ \color{red} Need a phone} \\
-$\phantom{^*}$Phone: +1 512 232 2459\\
+$\phantom{^{**}}$Phone: +1 512 232 2459\\
 
@@
 
+Need a section about:
+Rate4Site is completely different from the entropy variation and we do not know why. Explain the difference.
+
 \section{Acknowledgements}
@@
 
+Figures that are still needed...
+Table 1 - First make a new table with the mean values Entropy
+Table 2 - First make a new table with the mean values Rate4Site
+
+\begin{center}
+	\begin{tabular}{ | c | c | c |  }
+	\hline
+	  1 & 2 & 3 \\   \hline	
+	 4 & 5 & 6 \\ 	 \hline
+	  7 & 8 & 9 \\
+	  \hline
+	\end{tabular}
+\end{center}
+
+Need a two panel figure \\
+First plot the mean entropy vs the variance of entropy (then pick points that are in each of the four corners).
+This is B.
+I need to circle these by pointing an additional geom point and open circles and put numbers.
+\\
+Pick these proteins and then plot the distributions. This is A. \\
+Basically this figure will highlight the difference between mean entropy and variance in entropy. \\
+
+In the introduction and abstract need to better highlight the two papers (Yeh and Shah) that contradict on the surface and they do not. Hold off on Rate4Site for Now.
+
 \end{document}

### 2015-05-30 manuscript/manuscript.tex
Adding tables
Adding linear model table

@@
 \usepackage{color}
+\usepackage{array}
 
@@
 \begin{center}
-	\begin{tabular}{ | c | c | c |  }
+	\begin{table}[H]
+	\begin{tabular}{| p{1.5cm} | p{2.5cm} | p{2cm} | p{2cm} | }
 		\hline
-	  1 & 2 & 3 \\   \hline	
-	 4 & 5 & 6 \\ 	 \hline
-	  7 & 8 & 9 \\
+		Dataset & Mean $\rho$ Entropy-CN  & Mean $\rho$ Entropy-WCN & Mean $\rho$ Entropy-RSA \\
+		\hline	
+		Enzyme &  $-0.445$  & $-0.432$ & $0.464$ \\
+		\hline
+		Virus & $-0.213$ & $-0.217$ & $0.237$ \\
 		\hline	
 	\end{tabular}
+	\caption{NEED A CAPTION}
+	\label{table:entropy_stats}
+	\end{table}
+\end{center}
+
+\begin{center}
+	\begin{table}[H]
+	\begin{tabular}{| p{1.5cm} | p{2.5cm} | p{2cm} | p{2cm} |  }
+		\hline
+		Dataset & Mean $\rho$ Rate4Site-CN  & Mean $\rho$ Rate4Site-WCN & Mean $\rho$ Rate4Site-RSA \\
+		\hline	
+		Enzyme &  $-0.445$  & $-0.432$ & $0.464$ \\
+		\hline
+		Virus & $-0.213$ & $-0.217$ & $0.237$ \\
+		\hline	
+	\end{tabular}
+	\caption{NEED A CAPTION}0.07859
+	\label{table:entropy_stats}
+	\end{table}
+\end{center}
+
+%Should I report the intercept%
+\begin{center}
+	\begin{table}[H]
+	\begin{tabular}{ | p{5.0cm} | p{3cm} | p{3cm} | p{3cm} | }
+	\hline
+	Model & Coefficient x  & Coefficient y & Coefficient z \\	
+	\hline
+	CN-Entropy $\rho$ $ = $ Mean Entropy + Dataset + Dataset*Mean Entropy & $0.079 ^{**}$ & $0.382^{***}$  & $-0.583$ \\	
+	\hline
+	WCN-Entropy $\rho$ $ = $ Mean Entropy + Dataset + Dataset*Mean Entropy & $0.049$ & $0.323^{***}$ & $-0.568$  \\	
+	\hline
+	RSA-Entropy $\rho$ $ = $ Mean Entropy + Dataset + Dataset*Mean Entropy &  $-0.0300$ & $-0.336^{***}$ & $0.903^{*}$ \\
+
+	\hline
+	CN-Entropy $\rho$ $ = $ Variance Entropy + Dataset + Dataset*Variance Entropy & $-0.522^{***}$ &  $0.013$ & $-0.265$ \\	
+	\hline
+	WCN-Entropy $\rho$ $ = $ Variance Entropy + Dataset + Dataset*Variance Entropy & $-0.464^{***}$ &  $0.023$ & $-0.325$ \\	
+	\hline
+	RSA-Entropy $\rho$ $ = $ Variance Entropy + Dataset + Dataset*Variance Entropy & $0.415^{***}$ & $-0.076$ & $0.927$ \\
+
+	\hline
+	\end{tabular}
+	\caption{NEED A CAPTION. *** means p-value is less than 0.001. ** means p-value is less than 0.01.  * means p-value is less than 0.05. }
+	\label{table:rate_stats}
+	\end{table}
 \end{center}

### 2015-05-31 manuscript/manuscript.tex
Adding figures

@@
 However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.
-}
+In the introduction and abstract need to better highlight the two papers (Yeh and Shah) that contradict on the surface and they do not.
+Hold off on Rate4Site for Now.}
 \end{abstract}
@@
 
-Figures that are still needed...
-Table 1 - First make a new table with the mean values Entropy
-Table 2 - First make a new table with the mean values Rate4Site
+        \begin{figure}[H]
+            \centerline{\includegraphics[width=7.5in]{protein_ex_plot.pdf}}     
+            \caption{NEED A CAPTION!!!. Basically this figure will highlight the difference between mean entropy and variance in entropy.}
+            \label{fig:protein_ex_plot}
+    \end{figure}
 
@@
 	\end{tabular}
-	\caption{NEED A CAPTION}0.07859
+	\caption{NEED A CAPTION}
 	\label{table:entropy_stats}
@@
 	\hline
+
 	RSA-Entropy $\rho$ $ = $ Mean Entropy + Dataset + Dataset*Mean Entropy &  $-0.0300$ & $-0.336^{***}$ & $0.903^{*}$ \\
@@
 
-Need a two panel figure \\
-First plot the mean entropy vs the variance of entropy (then pick points that are in each of the four corners).
-This is B.
-I need to circle these by pointing an additional geom point and open circles and put numbers.
-\\
-Pick these proteins and then plot the distributions. This is A. \\
-Basically this figure will highlight the difference between mean entropy and variance in entropy. \\
-
-In the introduction and abstract need to better highlight the two papers (Yeh and Shah) that contradict on the surface and they do not. Hold off on Rate4Site for Now.
-
 \end{document}

### 2015-05-31 manuscript/manuscript.tex
Adding figure refs

@@
 On average the correlations between CN and entropy are larger in absolute magnitude in the enzyme proteins as compared to viral proteins.
-The average $\rho$ between each structural property and each measure of site variability (entropy and evolutionary rates) can be seen in {\color{red}Table 1 and Table 2}.
+The average $\rho$ between each structural property and each measure of site variability (entropy and evolutionary rates) can be seen in Table \ref{table:entropy_stats}  and Table \ref{table:rate_stats}.
 The viral proteins experience much lower site variability.
@@
 
-\indent  The average site variability of a given protein does not seem to be a significant determinant of the strength of structural correlations (Figure 1A, 2A, 3A).
+\indent  The average site variability of a given protein does not seem to be a significant determinant of the strength of structural correlations (Figures \ref{fig:entropy_cn_cor}A, \ref{fig:entropy_wcn_cor}A, \ref{fig:entropy_rsa_cor}A).
 However, when examining the variance of entropy there is a clear trend within the enzyme proteins.
-Proteins with a higher variance in site variability across the protein typically have have lower correlations (Figure 1B, 2B, 3B).
+Proteins with a higher variance in site variability across the protein typically have have lower correlations (Figures \ref{fig:entropy_cn_cor}B, \ref{fig:entropy_wcn_cor}B, \ref{fig:entropy_rsa_cor}B).
 If you extrapolate the trend from the enzyme proteins, the viral proteins follow a similar trend.
@@
             \caption{NEED A CAPTION!!!.}
-            \label{fig:seqent_structure_cors}
+            \label{fig:entropy_cn_cor}
     \end{figure}
@@
             \caption{NEED A CAPTION!!!.}
-            \label{fig:seqent_structure_cors}
+            \label{fig:entropy_wcn_cor}
     \end{figure}
@@
             \caption{NEED A CAPTION!!!.}
-            \label{fig:seqent_structure_cors}
+            \label{fig:entropy_rsa_cor}
     \end{figure}
@@
             \caption{NEED A CAPTION!!!.}
-            \label{fig:seqent_structure_cors}
+            \label{fig:rate_cor}
     \end{figure}
@@
             \caption{NEED A CAPTION!!!.}
-            \label{fig:seqent_structure_cors}
+            \label{fig:var_rate_rsa_cor}
     \end{figure}
@@
             \caption{NEED A CAPTION!!!.}
-            \label{fig:seqent_structure_cors}
+            \label{fig:var_entropy_rate_cor}
     \end{figure}
@@
 	\caption{NEED A CAPTION}
-	\label{table:entropy_stats}
+	\label{table:rate_stats}
 	\end{table}
@@
 	\caption{NEED A CAPTION. *** means p-value is less than 0.001. ** means p-value is less than 0.01.  * means p-value is less than 0.05. }
-	\label{table:rate_stats}
+	\label{table:model_stats}
 	\end{table}

### 2015-05-31 manuscript/manuscript.tex
Adding some preliminary figure captions

@@
 The correlation is negative between local packing density (both WCN and CN) and the variance of entropy is negative.
-Unlike with entropy, there is no relationship between variance of evolutionary rates at sites and any of the measured structural properties.
-There also is a wide spread int he variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
+Unlike with entropy, there is no relationship between variance of evolutionary rates at sites and any of the measured structural properties (Figure  .
+There also is a wide spread in the variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
+Although both evolutionary rate (as measured by Rate4Site) and entropy are measures of site variability, these quantities are distinctly different.
+Rate4Site measures the rate at which a site changes over time whereas site entropy measure the absolute variation at a site.
+A site may have a high evolutionary rate if it changes between frequently between the same few amino acids but have low entropy.
 
@@
 The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing High-performance computing resources.
-ELJ is is funded by a National Science Graduate Research Fellowship, grant number DGE-1110007.
+ELJ is funded by a National Science Graduate Research Fellowship, grant number DGE-1110007.
 COW is funded by {\color{red} Which grants??}.
 AS is funded by {\color{red} Which grants??}.
-....
-
-%\clearpage
-%\newpage
-
-%\cleardoublepage
 
@@
             \centerline{\includegraphics[width=7.5in]{entropy_cn_cor.pdf}}     
-            \caption{NEED A CAPTION!!!.}
+            \caption{Correlations between the Contact Number (CN) and Entropy.
+For each protein we calculate the Spearman Correlation coefficient between CN and entropy at each site within the protein.
+A) Comparison between the correlation coefficients and mean entropy of site in each protein.
+B) Comparison between correlation coefficients and the variance of entropy of each protein.
+The enzyme proteins are colored in black and the virus proteins are visualized in red.
+}
             \label{fig:entropy_cn_cor}
@@
             \centerline{\includegraphics[width=7.5in]{entropy_wcn_cor.pdf}}     
-            \caption{NEED A CAPTION!!!.}
+            \caption{Correlations between the Weighted Contact Number (WCN) and Entropy.
+A) Comparison between the correlation coefficients and mean entropy of site in each protein.
+B) Comparison between correlation coefficients and the variance of entropy of each protein.
+The enzyme proteins are colored in black and the virus proteins are visualized in red.}
             \label{fig:entropy_wcn_cor}
@@
             \centerline{\includegraphics[width=7.5in]{entropy_rsa_cor.pdf}}     
-            \caption{NEED A CAPTION!!!.}
+            \caption{Correlations between Relative Solvent Accessibility (RSA) and Entropy.
+A) Comparison between the correlation coefficients and mean entropy of site in each protein.
+B) Comparison between correlation coefficients and the variance of entropy of each protein.
+The enzyme proteins are colored in black and the virus proteins are visualized in red.}
             \label{fig:entropy_rsa_cor}
@@
             \centerline{\includegraphics[width=7.5in]{rate_cor.pdf}}     
-            \caption{NEED A CAPTION!!!.}
+            \caption{Comparison of Rate4Site correlations and the variance of Rate4Site at sites.
+A) Comparison between the correlation coefficients between CN and Rate4Site and variance of Rate4Site of site in each protein.
+B) Comparison between correlation coefficients between WCN and Rate4Site and the variance of entropy of each protein.
+The enzyme proteins are colored in black and the virus proteins are visualized in red.}
             \label{fig:rate_cor}
@@
             \centerline{\includegraphics[width=5.0in]{var_rate_rsa_cor.pdf}}     
-            \caption{NEED A CAPTION!!!.}
+            \caption{Comparison of the correlations between RSA and Rate4Site and the variance of Rate4Site at sites.
+The enzyme proteins are colored in black and the virus proteins are visualized in red.}
             \label{fig:var_rate_rsa_cor}
@@
             \centerline{\includegraphics[width=5.0in]{var_entropy_rate_cor.pdf}}     
-            \caption{NEED A CAPTION!!!.}
+            \caption{Relationship between WCN-Rate4Site correlations and the variance of entropy at sites. The enzyme proteins are colored in black and the virus proteins are colored red. }
             \label{fig:var_entropy_rate_cor}
@@
             \centerline{\includegraphics[width=7.5in]{protein_ex_plot.pdf}}     
-            \caption{NEED A CAPTION!!!. Basically this figure will highlight the difference between mean entropy and variance in entropy.}
+            \caption{Comparison of the mean entropy and the variance of entropy for individual proteins.
+A)  Variance in entropy at sites compared against overall mean entropy for proteins.
+Four proteins of interest are highlighted by open circles.
+The enzyme proteins are colored in black and the virus proteins are colored red B) Distributions of site-wise entropy values for four enzyme proteins.
+}
             \label{fig:protein_ex_plot}
@@
 	\end{tabular}
-	\caption{NEED A CAPTION}
+	\caption{Averages of the Spearman Correlation Coefficients between Structural Properties and Entropy.
+The structural properties analyzed are solvent accessibility as calculated by RSA and packing density as calculated by CN and WCN.}
 	\label{table:entropy_stats}
@@
 	\end{tabular}
-	\caption{NEED A CAPTION}
+	\caption{Averages of the Spearman Correlation Coefficients between Structural Properties and Evolutionary Rate.
+Evolutionary rate is calculated with Rate4Site.
+The structural properties analyzed are solvent accessibility as calculated by RSA and packing density as calculated by CN and WCN}
 	\label{table:rate_stats}
@@
 	\end{tabular}
-	\caption{NEED A CAPTION. *** means p-value is less than 0.001. ** means p-value is less than 0.01.  * means p-value is less than 0.05. }
+	\caption{Linear models predicting structural correlations with various quantities.
+Coefficient x represents the coefficient of the predictor, coefficient x is the coefficient of the second predictor and coefficient z is the coefficient of the third predictor in each model.
+*** means p-value is less than 0.001.
+** means p-value is less than 0.01.
+* means p-value is less than 0.05.
+}
 	\label{table:model_stats}

### 2015-05-31 manuscript/manuscript.tex
Adding results

@@
     \subsection*{Calculation of Structural Properties}
-In our analysis we used two types of measures of local packing density used in previous studies: Contact Numbe r(CN) and Weighted Contact Number(WCN).
+In our analysis we used two types of measures of local packing density used in previous studies: Contact Number (CN) and Weighted Contact Number(WCN).
 For the purposes of our comparison between the two datasets of interest we used the CN and WCN for the enzyme proteins calculated in \cite{Huangetal2014} and the CN and WCN values for the viral proteins from \cite{Shahmoradietal2014}.
@@
 The correlation is negative between local packing density (both WCN and CN) and the variance of entropy is negative.
-Unlike with entropy, there is no relationship between variance of evolutionary rates at sites and any of the measured structural properties (Figure  .
+Unlike with entropy, there is no relationship between variance of evolutionary rates at sites and any of the measured structural properties (Figures \ref{fig:var_rate_cor}, \ref{fig:var_rate_rsa_cor}).
 There also is a wide spread in the variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
@@
 A site may have a high evolutionary rate if it changes between frequently between the same few amino acids but have low entropy.
+It appears that the rate at which a site evolves is not important for predicting the strength of structure-sequence relationships.
+
+\indent In order to further examine the relationship between entropy and the structure-sequence relationships, we used mean entropy and variance of entropy as predictors of structure-sequence relationships.
+Table \ref{table:model_stats} illustrates the the coefficients of venous linear models.
+In all three models that use the mean entropy as a predictor of structure-entropy correlations, the dataset is a significant predictor.
+For WCN-Entropy correlations, mean entropy is not a significant predictor.
+Although for CN-Entropy correlations, mean entropy is significant, the coefficient is 0.079 and therefore predictive power is low.
+The mean site variability of a protein as measured by entropy is not indicative of whether packing density can be used to predict site variation.
+Mean Entropy is also not a significant predictor for the strength of the relationship between RSA and Entropy.
+For all linear models where mean entropy and dataset were used to predict structure-sequence correlations, dataset was a significant predictor.
+This means that enzyme and virus proteins have different correlations when used mean entropy as a predictor.
+This agrees with differences seen in the previous works by \cite{Yehetal2014} and \cite{Shahmoradietal2014}.
+
+\indent However, when looking at the variance in entropy there are some stark differences.
+For all structural predictors (i.e., CN, WCN and RSA), the variance in entropy at sites within a protein is a significant predictor of the strength of structure-sequence relations.
+For packing density, proteins with a larger variance in site-wise entropy have lower correlations.
+Proteins with a higher variance in entropy tend to have stronger RSA-Entropy correlations.
+When using variance and dataset as predictors of structure-sequence correlations, dataset was not a significant predictor in any model.
+When looking at the variance of entropy, proteins within both datasets at in a similar fashion.
+Overall trends seen between the variance of site variation and the strength of structure-sequence relationships is preserved across both datasets.
 
@@
 The enzyme proteins are colored in black and the virus proteins are visualized in red.}
-            \label{fig:rate_cor}
+            \label{fig:var_rate_cor}
     \end{figure}
@@
 
-        \begin{figure}[H]
-            \centerline{\includegraphics[width=5.0in]{var_entropy_rate_cor.pdf}}     
-            \caption{Relationship between WCN-Rate4Site correlations and the variance of entropy at sites. The enzyme proteins are colored in black and the virus proteins are colored red. }
-            \label{fig:var_entropy_rate_cor}
-    \end{figure}
+%       \begin{figure}[H]
+%            \centerline{\includegraphics[width=5.0in]{var_entropy_rate_cor.pdf}}     
+%            \caption{Relationship between WCN-Rate4Site correlations and the variance of entropy at sites. The enzyme proteins are colored in black and the virus proteins are colored red. }
+%            \label{fig:var_entropy_rate_cor}
+%    \end{figure}
 
@@
 Evolutionary rate is calculated with Rate4Site.
-The structural properties analyzed are solvent accessibility as calculated by RSA and packing density as calculated by CN and WCN}
+The structural properties analyzed are solvent accessibility as calculated by RSA and packing density as calculated by CN and WCN.}
 	\label{table:rate_stats}

### 2015-05-31 manuscript/manuscript.tex
Adding some more results

@@
 There also is a wide spread in the variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
-Although both evolutionary rate (as measured by Rate4Site) and entropy are measures of site variability, these quantities are distinctly different.
+
+\indent Although both evolutionary rate (as measured by Rate4Site) and entropy are measures of site variability, these quantities are distinctly different.
 Rate4Site measures the rate at which a site changes over time whereas site entropy measure the absolute variation at a site.
@@
 It appears that the rate at which a site evolves is not important for predicting the strength of structure-sequence relationships.
+However, the absolute amount of variation at a site can be used to predict the strength structure-sequence relationships.
+
+\indent For a given protein its mean entropy and the variance might be difference.
+A protein can have a high mean entropy but have a low variance and vice versa.
+Figure \ref{fig:protein_ex_plot}A details the relationship between the average entropy across sites within proteins and the variance of entropy across sites within a given protein.
+Additionally, the distribution of site-wise variance varies greatly between proteins even when they from the same dataset (Figure \ref{fig:protein_ex_plot}B).
+Therefore the mean entropy of a protein as a the variance in entropy at sites may be predictive for structure-sequence relationships.
 
@@
 {\color{red}SOME DISCUSSION.....} \\
-
-Need a section about:
-Rate4Site is completely different from the entropy variation and we do not know why. Explain the difference.
 

### 2015-05-31 manuscript/manuscript.tex
Adding an abstract

@@
 \begin{abstract}
-{\color{red}Recent work has shown that structural properties are capable of predicting site-specific sequence variability for a given protein.
+Recent work has shown that structural properties are capable of predicting site-specific sequence variability for a given protein.
 However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.
-In the introduction and abstract need to better highlight the two papers (Yeh and Shah) that contradict on the surface and they do not.
-Hold off on Rate4Site for Now.}
+Additionally there have been conflicting reports about about whether local packing density or solvent accessibility is the best predictor of site variability in proteins.
+According to \cite{Yehetal2014}, local packing density is the strongest predictor of site-wise variability.
+However, recently \cite{Shahmoradietal2014} compared local packing density and solvent accessibility in virus proteins and found that RSA is a stronger predictor of site-wise variability.
+Here we present research that suggests that differences in the correlations between these two datasets is due to differences in the site variability among proteins within in each dataset.
+Specifically proteins with a larger variance in entropy among sites exhibit stronger structure-sequence correlations between both local packing density and solvent accessibility.
 \end{abstract}

### 2015-06-01 manuscript/manuscript.tex
Adding a discussion

@@
 \label{sec:dcr}
-{\color{red}SOME DISCUSSION.....} \\
+Structure imposes constraints on the the evolution of proteins.
+Therefore structural predictors can be used to predict the variability at sites.
+Two main determinants of variability at protein sits are solvent accessibility and local packing density.
+Sites that are on surface of the protein tend to have higher solvent accessibility and exhibit more site variability.
+Sites that are densely packed and have more contacts tend to evolve slower and exhibit less sequence variability.
+However, it has not been well understood which of these two predictors, local packing density and solvent accessibility, is the main determinant of site variability in proteins.
+Two recent papers \citep{Shahmoradietal2014, Yehetal2014} give conflicting reports on which is the main determinant.
+In a study done on monomeric enzyme proteins, \cite{Yehetal2014} reported that local packing density, as measured by Contact Number and Weighted Contact Number, is the strongest determinant of site variability.
+\cite{Shahmoradietal2014} found that which a group of virus proteins, solvent accessibility is more predictive of site-wise variability.
+
+\indent Here we performed examined the relationship between site variability and the strength of structure-sequence relationships.
+We found that proteins with a larger variance in site variability as measured by sequence entropy, have stronger structure-sequence relationships on average.
+However, when looking at the mean entropy values of proteins there is a distinct difference between the virus and enzyme datasets.
+Overall the enzyme proteins have a higher site variability than the virus proteins.
+When predicting the structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variations.
+It appears that site variability and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships.
+Additionally the reason that the enzyme proteins in \cite{Yehetal2014} had higher correlation coefficients is that on average, the enzyme proteins have larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}.
+Moreover, since the variance is entropy is is a more significant predictor of correlation's in local packing density, an increase level of variance in site variability will have a larger effect on the local packing density correlations.
+This would explain why the local packing density correlations were of a higher magnitude compared to the RSA correlations within the enzyme proteins of /\cite{Yehetal2014}.
+Finally, The variance in evolutionary rate among sites of a protein does not have an effect on the correlation magnitude.
+This suggests that the absolute variability at sites is a better predictor of whether structural determinants can be used to predict variability at sites  as opposed to evolutionary rate.
+Therefore when examining structural determinants of site variability, entropy is a may be a more meaningful quantity.
 

### 2015-06-01 manuscript/manuscript.tex
Adding caption changes

@@
 The enzyme proteins are colored in black and the virus proteins are visualized in red.
+For both WCN and CN the viral proteins have lower mean entropy and variance of entropy.
+The line indicates a linear regression through the enzyme proteins with variance entropy as a single predictor of the spearman correlations.
 }
@@
 B) Comparison between correlation coefficients between WCN and Rate4Site and the variance of entropy of each protein.
-The enzyme proteins are colored in black and the virus proteins are visualized in red.}
+The enzyme proteins are colored in black and the virus proteins are visualized in red.
+There is no observable trend between the variance of evolutionary rate and the spearman correlations.
+}
             \label{fig:var_rate_cor}
@@
             \caption{Comparison of the correlations between RSA and Rate4Site and the variance of Rate4Site at sites.
-The enzyme proteins are colored in black and the virus proteins are visualized in red.}
+The enzyme proteins are colored in black and the virus proteins are visualized in red.
+The variance of evolutionary rate of a protein, as measured by Rate4Site, does not a correlation within the spearman correlation between RSA and evolutionary rate for a given protein.}
             \label{fig:var_rate_rsa_cor}

### 2015-06-01 manuscript/manuscript.tex
Adding manuscript changes

@@
 However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.
-Additionally there have been conflicting reports about about whether local packing density or solvent accessibility is the best predictor of site variability in proteins.
+Additionally, there have been conflicting reports about about whether local packing density or solvent accessibility is the best predictor of site variability in proteins.
 According to \cite{Yehetal2014}, local packing density is the strongest predictor of site-wise variability.
-However, recently \cite{Shahmoradietal2014} compared local packing density and solvent accessibility in virus proteins and found that RSA is a stronger predictor of site-wise variability.
-Here we present research that suggests that differences in the correlations between these two datasets is due to differences in the site variability among proteins within in each dataset.
+However, recently, \cite{Shahmoradietal2014} compared local packing density and solvent accessibility in viral proteins and found that relative solvent accessibility is a stronger predictor of site-wise variability.
+Here we present research that suggests that differences in the correlations between these two datasets are due to differences in the site variability among proteins within in each dataset.
 Specifically proteins with a larger variance in entropy among sites exhibit stronger structure-sequence correlations between both local packing density and solvent accessibility.
@@
 Proteins are subject to a number of biophysical and functional constraints \citep{Scherreretal2012, Wilkeetal2010}.
-These constraints result site-specific patterns of sequence variability within a protein.
+These constraints result in site-specific patterns of sequence variability within a protein.
 Recently several site-specific structural properties that can explain patterns of sequence variability in proteins have been identified.
-One of the earliest examples is Relative Site Accessibility (RSA).
-\cite{Fransozaetal2009} identified RSA as the strong predictor of evolutionary rate and found that residues that are buried in the core of proteins tend to be more conserved than exposed residues close to the surface of the protein.
-In their analysis they considered the ability of both RSA and various definitions of residue packing density to predict evolutionary rate.
+One of the earliest examples was Relative Site Accessibility (RSA).
+\cite{Fransozaetal2009} identified RSA as the strongest predictor of evolutionary rate and found that residues that are buried within the core of proteins tend to be more conserved than exposed residues close to the surface of the protein.
+In their analysis, they considered the ability of both RSA and various definitions of residue packing density to predict evolutionary rate.
 They found that RSA and evolutionary rate shared a significant linear relationship.
-Afterwards, several other works \citep{Ramseyetal2011, Scherreretal2012} also found that RSA as a very significant predictor of evolutionary rate and found this linear relationship as well.
+Afterwards, several other works \citep{Ramseyetal2011, Scherreretal2012} also found that RSA as a significant predictor of evolutionary rate and supported this linear relationship.
 However, these papers all have the same flaw.
-During the course of their analysis they binned the protein sites and average over all sites within a bin when looking at the trend of RSA.
-This process may have produce artifacts that account for this strong linear trend.
+During the course of their analysis they binned the protein sites and averaged over all sites within a bin when determining the trend of RSA.
+This process may have produced artifacts that account for this strong linear trend between RSA and evolutionary rate.
 \\
-\indent Recently, \cite{Yehetal2014} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by CN and WCN \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
-A year later, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.
+\indent Recently, \cite{Yehetal2014} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by Contact Number and Weighted Contact Number \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
+Soon Afterwards, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.
 In this analysis they found that RSA had the strongest correlation with site variability as opposed to local packing density.
-Additionally the effect seen between CN and WCN was of a much smaller magnitude as compared to \citep{Yehetal2014}.
+Additionally the effect seen between CN and WCN was of a much smaller magnitude as compared to \cite{Yehetal2014}.
 Therefore the relationship between local pacing density, RSA and site-specific measures of sequence variability is not well understood.
@@
 The viral proteins range from 122 - 557 residues in length and each structure is accompanied by a sequence alignment of up to 2362 homologous sequences.
-Sequence alignments for both datasets were constructed by aligning the amino-acid sequences using the alignment software MAFFT \citep{Katohetal2002, Katohetal2005}, specifying the auto flag to select the optimal algorithm for the given data set.
-The alignments were then used to calculate the site-specific measures of sequence variability for each individual protein in both datasets.
+Sequence alignments for both datasets were constructed by aligning the amino-acid sequences using the alignment software MAFFT \citep{Katohetal2002, Katohetal2005}, specifying the auto flag to select the optimal algorithm for the given dataset.
+The alignments were then used to calculate site-specific measures of sequence variability for each individual protein in both datasets.
 To do so, we relied on two independent methods of measuring sequence variability.
@@
     where $P_{ij}$ is the relative frequency of amino acid $j$ at position $i$ in the alignment.
-The sequence entropy is a measure of variability at each site.
-We also calculated a measure of site-specific evolutionary rate for each protein using software Rate4site.
-First the Maximum Likelihood phylogenetic trees were inferred with RAxML, using the LG substitution matrix and the CAT model of rate heterogeneity \citep{Stamatakis2006, Stamatakis2014}.
+Sequence entropy is a measure of variability at each site.
+We also calculated a measure of site-specific evolutionary rate for each protein using the software Rate4site.
+First Maximum Likelihood phylogenetic trees were inferred with RAxML, using the LG substitution matrix and the CAT model of rate heterogeneity \citep{Stamatakis2006, Stamatakis2014}.
 For each structure, we then used the respective sequence alignment and phylogenetic tree to infer site-specific substitution rates with Rate4Site using the empirical Bayesian method and the JTT model of sequence evolution \citep{Mayroseetal2004}.
@@
     \subsection*{Calculation of Structural Properties}
-In our analysis we used two types of measures of local packing density used in previous studies: Contact Number (CN) and Weighted Contact Number(WCN).
+In our analysis we used two types of measures of local packing density used in previous studies: Contact Number (CN) and Weighted Contact Number (WCN).
 For the purposes of our comparison between the two datasets of interest we used the CN and WCN for the enzyme proteins calculated in \cite{Huangetal2014} and the CN and WCN values for the viral proteins from \cite{Shahmoradietal2014}.
@@
 In this case, $r_0$ = 13 as in the previous papers.
-Weighed Contact Number for a residue, $i$, is defined as in (Liao et al 2005, Huang et al 2014), as:
+Weighed Contact Number for a residue, $i$, is defined as in \cite{Liaoetal2005} and \cite{Huangetal2014} as:
 
@@
 The viral proteins experience much lower site variability.
-Figure 1 shows the correlations between CN and entropy for the virus and enzyme proteins.
-Even though on average the correlations in the enzyme proteins are larger as reported, by \cite{Yehetal2014}, the virus proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations.
+Figure 1 shows the correlations between CN and entropy for the viral and enzyme proteins.
+Even though on average the correlations in the enzyme proteins are larger as reported by \cite{Yehetal2014}, the viral proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations.
 
-\indent  The average site variability of a given protein does not seem to be a significant determinant of the strength of structural correlations (Figures \ref{fig:entropy_cn_cor}A, \ref{fig:entropy_wcn_cor}A, \ref{fig:entropy_rsa_cor}A).
+\indent For a given protein its mean entropy and the variance might be different.
+A protein can have a high mean entropy but have a low variance and vice versa.
+Figure \ref{fig:protein_ex_plot}A details the relationship between the average entropy across sites within proteins and the variance of entropy across sites.
+Additionally, the distribution of variance varies greatly between proteins even when they from the same dataset (Figure \ref{fig:protein_ex_plot}B).
+Therefore the mean entropy of a protein as well as the variance in entropy at sites may be predictive for structure-sequence relationships.
+
+\indent  The average site entropy of a given protein does not seem to be a significant determinant of the strength of structural correlations (Figures \ref{fig:entropy_cn_cor}A, \ref{fig:entropy_wcn_cor}A, \ref{fig:entropy_rsa_cor}A).
 However, when examining the variance of entropy there is a clear trend within the enzyme proteins.
-Proteins with a higher variance in site variability across the protein typically have have lower correlations (Figures \ref{fig:entropy_cn_cor}B, \ref{fig:entropy_wcn_cor}B, \ref{fig:entropy_rsa_cor}B).
+Proteins with a higher variance in site variability across the protein typically have lower correlations (Figures \ref{fig:entropy_cn_cor}B, \ref{fig:entropy_wcn_cor}B, \ref{fig:entropy_rsa_cor}B).
 If you extrapolate the trend from the enzyme proteins, the viral proteins follow a similar trend.
 The correlation between the $\rho$ between RSA and entropy and the variance of entropy is positive.
-The correlation is negative between local packing density (both WCN and CN) and the variance of entropy is negative.
-Unlike with entropy, there is no relationship between variance of evolutionary rates at sites and any of the measured structural properties (Figures \ref{fig:var_rate_cor}, \ref{fig:var_rate_rsa_cor}).
+Proteins with a larger variance in site entropy have the strongest correlations.
+The correlation between local packing density (both WCN and CN) and the variance of entropy is negative.
+Unlike entropy, there is no relationship between the variance of evolutionary rates at sites and any of the measured structural properties (Figures \ref{fig:var_rate_cor}, \ref{fig:var_rate_rsa_cor}).
 There also is a wide spread in the variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
 
-\indent Although both evolutionary rate (as measured by Rate4Site) and entropy are measures of site variability, these quantities are distinctly different.
+\indent Although both evolutionary rate, as measured by Rate4Sit,e and entropy are measures of site variability, these quantities are distinctly different.
 Rate4Site measures the rate at which a site changes over time whereas site entropy measure the absolute variation at a site.
-A site may have a high evolutionary rate if it changes between frequently between the same few amino acids but have low entropy.
+A site may have a high evolutionary rate if it changes  frequently between the same few amino acids but have low entropy.
 It appears that the rate at which a site evolves is not important for predicting the strength of structure-sequence relationships.
-However, the absolute amount of variation at a site can be used to predict the strength structure-sequence relationships.
+However, the absolute amount of variation at a site can be used to predict the strength of structure-sequence relationships.
 
-\indent For a given protein its mean entropy and the variance might be difference.
-A protein can have a high mean entropy but have a low variance and vice versa.
-Figure \ref{fig:protein_ex_plot}A details the relationship between the average entropy across sites within proteins and the variance of entropy across sites within a given protein.
-Additionally, the distribution of site-wise variance varies greatly between proteins even when they from the same dataset (Figure \ref{fig:protein_ex_plot}B).
-Therefore the mean entropy of a protein as a the variance in entropy at sites may be predictive for structure-sequence relationships.
-
-\indent In order to further examine the relationship between entropy and the structure-sequence relationships, we used mean entropy and variance of entropy as predictors of structure-sequence relationships.
-Table \ref{table:model_stats} illustrates the the coefficients of venous linear models.
-In all three models that use the mean entropy as a predictor of structure-entropy correlations, the dataset is a significant predictor.
+\indent In order to further examine the relationship between entropy and the structure-sequence relationships, we used the mean entropy and variance of entropy at sites as predictors of structure-sequence relationships.
+Table \ref{table:model_stats} illustrates the coefficients of various linear models.
 For WCN-Entropy correlations, mean entropy is not a significant predictor.
-Although for CN-Entropy correlations, mean entropy is significant, the coefficient is 0.079 and therefore predictive power is low.
-The mean site variability of a protein as measured by entropy is not indicative of whether packing density can be used to predict site variation.
+Although for CN-Entropy correlations, mean entropy is significant the coefficient is 0.079 and therefore predictive power is low.
+Therefore the mean site variability of a protein as measured by entropy is not indicative of whether packing density can be using to predict site variation.
 Mean Entropy is also not a significant predictor for the strength of the relationship between RSA and Entropy.
-For all linear models where mean entropy and dataset were used to predict structure-sequence correlations, dataset was a significant predictor.
-This means that enzyme and virus proteins have different correlations when used mean entropy as a predictor.
+For all linear models where mean entropy and dataset were used to predict structure-sequence correlations, dataset was a significant predictor (Table \ref{table:model_stats}).
+This means that enzyme and virus proteins have different correlations when using mean entropy as a predictor.
 This agrees with differences seen in the previous works by \cite{Yehetal2014} and \cite{Shahmoradietal2014}.
@@
 When using variance and dataset as predictors of structure-sequence correlations, dataset was not a significant predictor in any model.
-When looking at the variance of entropy, proteins within both datasets at in a similar fashion.
-Overall trends seen between the variance of site variation and the strength of structure-sequence relationships is preserved across both datasets.
+When looking at the variance of entropy, proteins within both datasets at in a similar fashion and overall trends seen between the variance of site variation and the strength of structure-sequence relationships is preserved across both datasets.
 
@@
 \label{sec:dcr}
-Structure imposes constraints on the the evolution of proteins.
+Structure imposes constraints on the evolution of proteins.
 Therefore structural predictors can be used to predict the variability at sites.
-Two main determinants of variability at protein sits are solvent accessibility and local packing density.
+Two main determinants of variability at protein sites are solvent accessibility and local packing density.
 Sites that are on surface of the protein tend to have higher solvent accessibility and exhibit more site variability.
@@
 However, it has not been well understood which of these two predictors, local packing density and solvent accessibility, is the main determinant of site variability in proteins.
-Two recent papers \citep{Shahmoradietal2014, Yehetal2014} give conflicting reports on which is the main determinant.
+Two recent papers \citep{Shahmoradietal2014, Yehetal2014} give conflicting reports on which is the best predictor.
 In a study done on monomeric enzyme proteins, \cite{Yehetal2014} reported that local packing density, as measured by Contact Number and Weighted Contact Number, is the strongest determinant of site variability.
-\cite{Shahmoradietal2014} found that which a group of virus proteins, solvent accessibility is more predictive of site-wise variability.
+\cite{Shahmoradietal2014} performed a similar analysis on a series of viral proteins and found that solvent accessibility is more predictive of site variability.
 
-\indent Here we performed examined the relationship between site variability and the strength of structure-sequence relationships.
+\indent Here we examined the relationship between site variability and the strength of structure-sequence relationships.
 We found that proteins with a larger variance in site variability as measured by sequence entropy, have stronger structure-sequence relationships on average.
 However, when looking at the mean entropy values of proteins there is a distinct difference between the virus and enzyme datasets.
-Overall the enzyme proteins have a higher site variability than the virus proteins.
-When predicting the structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variations.
-It appears that site variability and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships.
+Overall the enzyme proteins have more site variability than the viral proteins.When predicting structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variability.
 Additionally the reason that the enzyme proteins in \cite{Yehetal2014} had higher correlation coefficients is that on average, the enzyme proteins have larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}.
-Moreover, since the variance is entropy is is a more significant predictor of correlation's in local packing density, an increase level of variance in site variability will have a larger effect on the local packing density correlations.
-This would explain why the local packing density correlations were of a higher magnitude compared to the RSA correlations within the enzyme proteins of /\cite{Yehetal2014}.
-Finally, The variance in evolutionary rate among sites of a protein does not have an effect on the correlation magnitude.
+Moreover, since the variance in entropy is a more significant predictor of correlations in local packing density, an increase level of variance in site variability will have a larger effect on the local packing density correlations.
+This would explain why the local packing density correlations were of a higher magnitude compared to the RSA correlations within the enzyme proteins of \cite{Yehetal2014}.
+It appears that site entropy and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships.
+Finally, the variance in evolutionary rate among sites of a protein does not have an effect on the correlation magnitude.
 This suggests that the absolute variability at sites is a better predictor of whether structural determinants can be used to predict variability at sites  as opposed to evolutionary rate.
@@
 \section{Acknowledgements}
-The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing High-performance computing resources.
+The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing high-performance computing resources.
 ELJ is funded by a National Science Graduate Research Fellowship, grant number DGE-1110007.
@@
             \centerline{\includegraphics[width=7.5in]{entropy_cn_cor.pdf}}     
-            \caption{Correlations between the Contact Number (CN) and Entropy.
+            \caption{Correlations between Contact Number (CN) and Entropy.
 For each protein we calculate the Spearman Correlation coefficient between CN and entropy at each site within the protein.
 A) Comparison between the correlation coefficients and mean entropy of site in each protein.
+The line indicates a linear regression through the enzyme proteins with mean entropy as a single predictor of the spearman correlations.
 B) Comparison between correlation coefficients and the variance of entropy of each protein.
@@
 The line indicates a linear regression through the enzyme proteins with variance entropy as a single predictor of the spearman correlations.
+Proteins with a larger variance in site entropy among sites tend to smaller CN-Entropy correlations in magnitude.
 }
@@
             \centerline{\includegraphics[width=7.5in]{entropy_wcn_cor.pdf}}     
-            \caption{Correlations between the Weighted Contact Number (WCN) and Entropy.
+            \caption{Correlations between Weighted Contact Number (WCN) and Entropy.
 A) Comparison between the correlation coefficients and mean entropy of site in each protein.
 B) Comparison between correlation coefficients and the variance of entropy of each protein.
-The enzyme proteins are colored in black and the virus proteins are visualized in red.}
+The enzyme proteins are colored in black and the virus proteins are visualized in red.
+Proteins with a larger variance in site entropy have smaller WCN-Entropy correlations in magnitude.}
             \label{fig:entropy_wcn_cor}
@@
 B) Comparison between correlation coefficients and the variance of entropy of each protein.
-The enzyme proteins are colored in black and the virus proteins are visualized in red.}
+The enzyme proteins are colored in black and the virus proteins are visualized in red.
+Proteins have a larger variance in site entropy have larger RSA-Entropy correlations.}
             \label{fig:entropy_rsa_cor}
@@
 The enzyme proteins are colored in black and the virus proteins are visualized in red.
-The variance of evolutionary rate of a protein, as measured by Rate4Site, does not a correlation within the spearman correlation between RSA and evolutionary rate for a given protein.}
+The variance of evolutionary rate of a protein, as measured by Rate4Site, does not have a correlation with the spearman correlation between RSA and evolutionary rate for a given protein.}
             \label{fig:var_rate_rsa_cor}
@@
 Four proteins of interest are highlighted by open circles.
-The enzyme proteins are colored in black and the virus proteins are colored red B) Distributions of site-wise entropy values for four enzyme proteins.
-}
+The enzyme proteins are colored in black and the virus proteins are colored red B) Distributions of site-wise entropy values for the four enzyme proteins highlighted in A.
+There are a variety of distributions in site entropy for different proteins.}
             \label{fig:protein_ex_plot}
@@
 	\caption{Linear models predicting structural correlations with various quantities.
-Coefficient x represents the coefficient of the predictor, coefficient x is the coefficient of the second predictor and coefficient z is the coefficient of the third predictor in each model.
-*** means p-value is less than 0.001.
-** means p-value is less than 0.01.
-* means p-value is less than 0.05.
+Coefficient x represents the coefficient of the first predictor, coefficient y is the coefficient of the second predictor and coefficient z is the coefficient of the third predictor in each model.
+*** means the p-value is less than 0.001.
+** means the p-value is less than 0.01.
+* means the p-value is less than 0.05.
 }

### 2015-06-01 scripts/create_figures.R

@@ -5,14 +5,22 @@ library(cowplot)
 correlation_data = read.csv("../data_summaries/output_files/all_protein_summary_stats.csv")
 correlation_data = rename(correlation_data, dataset = data_id)
 enzyme_data = filter(correlation_data, dataset == "enzyme")
+selected_data = all_data %>% select(pdb_id, dataset, entropy) %>% filter(pdb_id == "1G24" | pdb_id == "1DDJ" | pdb_id == "1IU4" | pdb_id == "1HQC")
+selected_data = rename(selected_data, PDB = pdb_id)
+
+selected_cor_data = correlation_data %>% filter(pdb_id == "1G24" | pdb_id == "1DDJ" | pdb_id == "1IU4" | pdb_id == "1HQC")
+
+
+all_data = read.csv("../data_summaries/output_files/all_pdb_data.csv")
+all_data= rename(all_data, dataset = data_id)
 
 p1 = ggplot(correlation_data, aes(x = mean_entropy, y = rsa_entropy_rho, color = dataset)) +
 	theme_classic() + 
   	xlab("Mean Entropy") + 
   	ylab("Correlation RSA-Entropy") + 
   	theme(text = element_text(size = 18)) +  #Get rid of bold
-  	theme(axis.line = element_line(size = 1.25)) + #Make this thinner
-  	theme(axis.ticks = element_line(size = 1.25)) + 
+  	theme(axis.line = element_line(size = 1.05)) + #Make this thinner
+  	theme(axis.ticks = element_line(size = 1.05)) + 
   	geom_point(size = 2) + 
   	stat_smooth(method = "lm", se = F, data = enzyme_data) + 
   	scale_color_manual(values = c("black", "red"))
@@ -24,14 +32,12 @@ p2 = ggplot(correlation_data, aes(x = var_entropy, y = rsa_entropy_rho, color = 
   	xlab("Variance Entropy") + 
   	ylab("Correlation RSA-Entropy") + 
   	 theme(text = element_text(size = 18)) + 
-  	theme(axis.text.x = element_text(face="bold")) + 
-  	theme(axis.line = element_line(size = 1.25)) + 
-  	theme(axis.ticks = element_line(size = 1.25)) + 
+  	theme(axis.line = element_line(size = 1.05)) + 
+  	theme(axis.ticks = element_line(size = 1.05)) + 
   	geom_point(size = 2) +
[…]

### 2015-06-01 scripts/linear_models.R

@@ -1,0 +1,20 @@
+library(dplyr)
+
+correlation_data = read.csv("../data_summaries/output_files/all_protein_summary_stats.csv")
+correlation_data = rename(correlation_data, dataset = data_id)
+
+model1 = lm(data = correlation_data, cn_entropy_rho ~ mean_entropy + dataset + dataset:mean_entropy)
+summary(model1)
+model2 = lm(data = correlation_data, wcn_entropy_rho ~ mean_entropy + dataset + dataset:mean_entropy)
+summary(model2)
+model3 = lm(data = correlation_data, rsa_entropy_rho ~ mean_entropy + dataset + dataset:mean_entropy)
+summary(model3)
+model4 = lm(data = correlation_data, cn_entropy_rho ~ var_entropy + dataset + dataset:var_entropy)
+summary(model4)
+model5 = lm(data = correlation_data, wcn_entropy_rho ~ var_entropy + dataset + dataset:var_entropy)
+summary(model5)
+model6 = lm(data = correlation_data, rsa_entropy_rho ~ var_entropy + dataset + dataset:var_entropy)
+summary(model6)
+
+mean_data = correlation_data %>% group_by(dataset) %>% summarize(mean_mean_entropy = mean(mean_entropy), sd_mean_entropy = sd(mean_entropy), var_mean_entropy = var(mean_entropy), mean_cn_entropy_rho = mean(cn_entropy_rho),  mean_wcn_entropy_rho = mean(wcn_entropy_rho), mean_rsa_entropy_rho = mean(rsa_entropy_rho), mean_rate_cn_rho = mean(rate_cn_rho), mean_rate_wcn_rho = mean(rate_wcn_rho), mean_rate_rsa_rho = mean(rate_rsa_rho))
+write.csv(mean_data, "../data_summaries/output_files/mean_summary_data.csv")

### 2015-06-02 manuscript/manuscript.tex
Making changes to figures

@@
             \begin{figure}[H]
-            \centerline{\includegraphics[width=5.0in]{var_rate_rsa_cor.pdf}}     
+            \centerline{\includegraphics[width=4in]{var_rate_rsa_cor.pdf}}     
             \caption{Comparison of the correlations between RSA and Rate4Site and the variance of Rate4Site at sites.
@@
 Four proteins of interest are highlighted by open circles.
-The enzyme proteins are colored in black and the virus proteins are colored red B) Distributions of site-wise entropy values for the four enzyme proteins highlighted in A.
+The enzyme proteins are colored in black and the virus proteins are colored red B-F) Distributions of site-wise entropy values for the five enzyme proteins highlighted in A.
 There are a variety of distributions in site entropy for different proteins.}
@@
 		\hline	
-		Enzyme &  $-0.445$  & $-0.432$ & $0.464$ \\
-		\hline
+		Enzyme &  $-0.445$  & $-0.432$ & $0.464$ \\[0.10cm]
+
 		Virus & $-0.213$ & $-0.217$ & $0.237$ \\
@@
 	\begin{table}[H]
-	\begin{tabular}{ | p{5.0cm} | p{3cm} | p{3cm} | p{3cm} | }
+	\begin{tabular}{ | p{5.0cm} | p{3cm} | p{3cm} | p{4cm} | }
 	\hline
-	Model & Coefficient x  & Coefficient y & Coefficient z \\	
+	Model & $\left \langle H \right \rangle$  & Dataset & Dataset*$\left \langle H \right \rangle$ \\	
 	\hline
-	CN-Entropy $\rho$ $ = $ Mean Entropy + Dataset + Dataset*Mean Entropy & $0.079 ^{**}$ & $0.382^{***}$  & $-0.583$ \\	
-	\hline
-	WCN-Entropy $\rho$ $ = $ Mean Entropy + Dataset + Dataset*Mean Entropy & $0.049$ & $0.323^{***}$ & $-0.568$  \\	
-	\hline
+	$\rho  \text{\_CN--H}$   $ = $  $\left \langle H \right \rangle$ + Dataset + Dataset*$\left \langle H \right \rangle$ & $0.079 ^{**}$ & $0.382^{***}$  & $-0.583$   \\[1.2cm]
 
-	RSA-Entropy $\rho$ $ = $ Mean Entropy + Dataset + Dataset*Mean Entropy &  $-0.0300$ & $-0.336^{***}$ & $0.903^{*}$ \\
+	$\rho  \text{\_WCN--H}$  $ = $  $\left \langle H \right \rangle$ + Dataset + Dataset*$\left \langle H \right \rangle$& $0.049$ & $0.323^{***}$ & $-0.568$  \\[1.2cm]
+
+	$\rho  \text{\_RSA--H}$  $ = $ $\left \langle H \right \rangle$ + Dataset + Dataset*$\left \langle H \right \rangle$ &  $-0.030$ & $-0.336^{***}$ & $0.903^{*}$ \\[1.2cm]
+
+	$\rho  \text{\_CN--H}$   $ = $  $\left \langle H \right \rangle$ + Dataset + Dataset*$\left \langle H \right \rangle$ & $0.077^{**}$ & $0.336^{***}$  & NA  \\[1.2cm]
+
+	$\rho  \text{\_WCN--H}$  $ = $  $\left \langle H \right \rangle$ + Dataset + Dataset*$\left \langle H \right \rangle$& $ 0.047$ & $0.278^{***}$ & NA  \\[1.2cm]
+
+	$\rho  \text{\_RSA--H}$  $ = $ $\left \langle H \right \rangle$ + Dataset + Dataset*$\left \langle H \right \rangle$ &  $-0.027$ & $ -0.264^{***}$ & NA \\
 
 	\hline
-	CN-Entropy $\rho$ $ = $ Variance Entropy + Dataset + Dataset*Variance Entropy & $-0.522^{***}$ &  $0.013$ & $-0.265$ \\	
+	\end{tabular}
+	\caption{Linear models predicting structural correlations with various quantities.
+Coefficient x represents the coefficient of the first predictor, coefficient y is the coefficient of the second predictor and coefficient z is the coefficient of the third predictor in each model.
+*** means the p-value is less than 0.001.
+** means the p-value is less than 0.01.
+* means the p-value is less than 0.05.
+}
+	\label{table:model_stats}
+	\end{table}
+\end{center}
+
+\begin{center}
+	\begin{table}[H]
+	\begin{tabular}{ | p{5.0cm} | p{3cm} | p{3cm} | p{4cm} | }
 	\hline
-	WCN-Entropy $\rho$ $ = $ Variance Entropy + Dataset + Dataset*Variance Entropy & $-0.464^{***}$ &  $0.023$ & $-0.325$ \\	
+	Model & Variance H  & Dataset & Dataset*Variance H \\	
 	\hline
-	RSA-Entropy $\rho$ $ = $ Variance Entropy + Dataset + Dataset*Variance Entropy & $0.415^{***}$ & $-0.076$ & $0.927$ \\
+	$\rho  \text{\_CN--H}$ $ = $ Variance H + Dataset + Dataset*Variance H & $-0.522^{***}$ &  $0.013$ & $-0.265$ \\[1.3cm]
+
+	$\rho  \text{\_WCN--H}$ $ = $ Variance H + Dataset + Dataset*Variance H& $-0.464^{***}$ &  $0.023$ & $-0.325$ \\ [1.3cm]
+
+	$\rho  \text{\_RSA--H}$ $ = $ Variance H + Dataset + Dataset*Variance H & $0.415^{***}$ & $-0.076$ & $0.927$ \\
+
+		$\rho  \text{\_CN--H}$ $ = $ Variance H + Dataset + Dataset*Variance H & $-0.524^{***}$ &  $0.004$ & NA \\[1.3cm]
+
+	$\rho  \text{\_WCN--H}$ $ = $ Variance H + Dataset + Dataset*Variance H& $-0.466 ^{***}$ &  $ 0.012$ & NA \\ [1.3cm]
+
+	$\rho  \text{\_RSA--H}$ $ = $ Variance H + Dataset + Dataset*Variance H & $0.422^{***}$ & $-0.044$ & NA \\
 

### 2015-06-03 manuscript/manuscript.tex
Addressed Amir's comments

@@
 However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.
-Additionally, there have been conflicting reports about about whether local packing density or solvent accessibility is the best predictor of site variability in proteins.
+There have two works that have investigated structural predictors of site variability which both present different results based on the strength of correlations of structural predictors.
 According to \cite{Yehetal2014}, local packing density is the strongest predictor of site-wise variability.
@@
 They found that RSA and evolutionary rate shared a significant linear relationship.
-Afterwards, several other works \citep{Ramseyetal2011, Scherreretal2012} also found that RSA as a significant predictor of evolutionary rate and supported this linear relationship.
+Afterwards, several other works  also found that RSA as a significant predictor of evolutionary rate and supported this linear relationship \citep{Ramseyetal2011, Scherreretal2012}.
 However, these papers all have the same flaw.
@@
 \indent Recently, \cite{Yehetal2014} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by Contact Number and Weighted Contact Number \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
-Soon Afterwards, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.
-In this analysis they found that RSA had the strongest correlation with site variability as opposed to local packing density.
-Additionally the effect seen between CN and WCN was of a much smaller magnitude as compared to \cite{Yehetal2014}.
+Soon Afterwards, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.{\color{blue} In this analysis they found that RSA had the strongest correlation with site variability as opposed to local packing density.} Additionally the effect seen between CN and WCN was of a much smaller magnitude as compared to \cite{Yehetal2014}.
 Therefore the relationship between local pacing density, RSA and site-specific measures of sequence variability is not well understood.
@@
 However, when examining the variance of entropy there is a clear trend within the enzyme proteins.
-Proteins with a higher variance in site variability across the protein typically have lower correlations (Figures \ref{fig:entropy_cn_cor}B, \ref{fig:entropy_wcn_cor}B, \ref{fig:entropy_rsa_cor}B).
+Proteins with a higher variance in site variability across the protein typically higher  correlations in magnitude (Figures \ref{fig:entropy_cn_cor}B, \ref{fig:entropy_wcn_cor}B, \ref{fig:entropy_rsa_cor}B).
 If you extrapolate the trend from the enzyme proteins, the viral proteins follow a similar trend.
@@
 
-\indent In order to further examine the relationship between entropy and the structure-sequence relationships, we used the mean entropy and variance of entropy at sites as predictors of structure-sequence relationships.
+\indent In order to further examine the relationship between entropy and the structure-sequence relationships, we used the mean entropy and variance of entropy at sites as predictors of the strength of structure-sequence relationships.
 Table \ref{table:model_stats} illustrates the coefficients of various linear models.
@@
 Although for CN-Entropy correlations, mean entropy is significant the coefficient is 0.079 and therefore predictive power is low.
-Therefore the mean site variability of a protein as measured by entropy is not indicative of whether packing density can be using to predict site variation.
 Mean Entropy is also not a significant predictor for the strength of the relationship between RSA and Entropy.
@@
 For all structural predictors (i.e., CN, WCN and RSA), the variance in entropy at sites within a protein is a significant predictor of the strength of structure-sequence relations.
-For packing density, proteins with a larger variance in site-wise entropy have lower correlations.
+For packing density, proteins with a larger variance in site-wise entropy have more negative correlations that are higher in magnitude.
 Proteins with a higher variance in entropy tend to have stronger RSA-Entropy correlations.
@@
 Sites that are densely packed and have more contacts tend to evolve slower and exhibit less sequence variability.
-However, it has not been well understood which of these two predictors, local packing density and solvent accessibility, is the main determinant of site variability in proteins.
+{\color{blue} However, it has not been well understood which of these two predictors, local packing density and solvent accessibility, is the main determinant of site variability in proteins.
 Two recent papers \citep{Shahmoradietal2014, Yehetal2014} give conflicting reports on which is the best predictor.
@@
 \cite{Shahmoradietal2014} performed a similar analysis on a series of viral proteins and found that solvent accessibility is more predictive of site variability.
+AMIR THINKS THIS IS NOT RELEVANT!}
 
@@
 Additionally the reason that the enzyme proteins in \cite{Yehetal2014} had higher correlation coefficients is that on average, the enzyme proteins have larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}.
-Moreover, since the variance in entropy is a more significant predictor of correlations in local packing density, an increase level of variance in site variability will have a larger effect on the local packing density correlations.
-This would explain why the local packing density correlations were of a higher magnitude compared to the RSA correlations within the enzyme proteins of \cite{Yehetal2014}.
+Moreover, since the variance in entropy is a more significant predictor of correlations in local packing density, an increased level of variance in site variability will have a larger effect on the local packing density correlations.
+{\color{blue} This would explain why the local packing density correlations were of a higher magnitude compared to the RSA correlations within the enzyme proteins of \cite{Yehetal2014}.
+THIS NEEDS TO BE EXPLAINED}.
 It appears that site entropy and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships.
-Finally, the variance in evolutionary rate among sites of a protein does not have an effect on the correlation magnitude.
+Finally, the variance in evolutionary rate among sites of a protein has a smaller an effect on the correlation magnitude as compared to entropy.
 This suggests that the absolute variability at sites is a better predictor of whether structural determinants can be used to predict variability at sites  as opposed to evolutionary rate.
 Therefore when examining structural determinants of site variability, entropy is a may be a more meaningful quantity.
+
+{\color{red} Amir's comment: I think it maybe better to not put so much emphasis on variance of entropy, but its implications and meaning.
+For example entropy variance is an indicator of sequence divergence, to some extent.}
 

### 2015-06-03 manuscript/manuscript.tex
Finished addressing Amir's comments

@@
 There have two works that have investigated structural predictors of site variability which both present different results based on the strength of correlations of structural predictors.
-According to \cite{Yehetal2014}, local packing density is the strongest predictor of site-wise variability.
-However, recently, \cite{Shahmoradietal2014} compared local packing density and solvent accessibility in viral proteins and found that relative solvent accessibility is a stronger predictor of site-wise variability.
+According to \cite{Yehetal2014}, both solvent accessibility and  local packing density are predictors of site-wise variability with local packing density being strongest predictor of site-wise variability.
+However, recently, \cite{Shahmoradietal2014} compared local packing density, residue flexibility, and solvent accessibility in viral proteins and found that relative solvent accessibility is a stronger predictor of site-wise variability.
+In addition, the strength of correlations with structural predictors were weaker in \cite{Shahmoradietal2014}.
 Here we present research that suggests that differences in the correlations between these two datasets are due to differences in the site variability among proteins within in each dataset.
@@
 \indent Recently, \cite{Yehetal2014} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by Contact Number and Weighted Contact Number \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
-Soon Afterwards, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.{\color{blue} In this analysis they found that RSA had the strongest correlation with site variability as opposed to local packing density.} Additionally the effect seen between CN and WCN was of a much smaller magnitude as compared to \cite{Yehetal2014}.
-Therefore the relationship between local pacing density, RSA and site-specific measures of sequence variability is not well understood.
+Soon Afterwards, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.
+In this analysis they found that RSA had a slightly stronger correlation with site variability as opposed to local packing density.
+Moreover, the effect seen between CN and WCN was of a much smaller magnitude as compared to \cite{Yehetal2014}.
+It is not well understood the reasoning behind the difference in magnitude in the correlations seen between the two studies.
 Here we attempt to reconcile the work done in this area.
@@
 
-\indent Although both evolutionary rate, as measured by Rate4Sit,e and entropy are measures of site variability, these quantities are distinctly different.
+\indent Although both evolutionary rate, as measured by Rate4Site and entropy are measures of site variability, these quantities are distinctly different.
 Rate4Site measures the rate at which a site changes over time whereas site entropy measure the absolute variation at a site.
@@
 Sites that are densely packed and have more contacts tend to evolve slower and exhibit less sequence variability.
-{\color{blue} However, it has not been well understood which of these two predictors, local packing density and solvent accessibility, is the main determinant of site variability in proteins.
-Two recent papers \citep{Shahmoradietal2014, Yehetal2014} give conflicting reports on which is the best predictor.
-In a study done on monomeric enzyme proteins, \cite{Yehetal2014} reported that local packing density, as measured by Contact Number and Weighted Contact Number, is the strongest determinant of site variability.
-\cite{Shahmoradietal2014} performed a similar analysis on a series of viral proteins and found that solvent accessibility is more predictive of site variability.
-AMIR THINKS THIS IS NOT RELEVANT!}
 
@@
 Moreover, since the variance in entropy is a more significant predictor of correlations in local packing density, an increased level of variance in site variability will have a larger effect on the local packing density correlations.
-{\color{blue} This would explain why the local packing density correlations were of a higher magnitude compared to the RSA correlations within the enzyme proteins of \cite{Yehetal2014}.
-THIS NEEDS TO BE EXPLAINED}.
 It appears that site entropy and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships.
@@
 Therefore when examining structural determinants of site variability, entropy is a may be a more meaningful quantity.
+
+Maybe change the focus: forget trying to figure out which is the best predictor: RSA or local packing density.
+Focus on bridging the gap on why there are differences in correlation strength.
+They had stronger correlations in the Yeh et al (2014) paper as compared to the Shahmoradi paper.
+Make a table of the mean correlations and variance from both papers.
+
+Proteins are under a number of constraints for function (Liberles et al Paper).
+Biophysical considerations that have been shown to affect function include: selection for protein stability, specificity for protein binding and selection against extraneous binding.
+Recently there have been several works that have looked into the effect of protein structure on protein function as well (Xia, Yeh, Shahmoradi, Echave, Huang)
 
@@
         \begin{figure}[H]
-            \centerline{\includegraphics[width=7.5in]{protein_ex_plot.pdf}}     
+            \centerline{\includegraphics[width=6.5in]{protein_ex_plot.pdf}}     
             \caption{Comparison of the mean entropy and the variance of entropy for individual proteins.

### 2015-06-09 manuscript/manuscript.tex
Adding citations to the manuscript

@@
 \usepackage[margin = 1.0 in]{geometry}
+%\bibliographystyle{ynt}
 \usepackage{natbib}
+%\usepackage[natbib=true,style=numeric,sorting=ynt]{biblatex}
 
@@
 \title{Variation in site entropy explains differences in structure-sequence relationships of proteins}
-\author{Eleisha L. Jackson$^{2^*}$, Amir Shahmoradi$^{1*}$, Claus O. Wilke$^{2^{**}}$}
+\author{Eleisha L. Jackson$^{1^*}$, Amir Shahmoradi$^{1*}$, Claus O. Wilke$^{1^{**}}$}
 \begin{document}
@@
 \noindent
-$^1$ Department of Physics, The University of Texas at Austin, Austin, TX 78712, USA \\
-$^2$ Institute of Cellular and Molecular Biology, Center for Computational Biology and Bioinformatics, and Department of Integrative Biology, The University of Texas at Austin, Austin, Texas, 78712 USA\\
+%$^1$ Department of Physics, The University of Texas at Austin, Austin, TX 78712, USA \\
+$^1$ Institute of Cellular and Molecular Biology, Center for Computational Biology and Bioinformatics, and Department of Integrative Biology, The University of Texas at Austin, Austin, Texas, 78712 USA\\
 
@@
 There have two works that have investigated structural predictors of site variability which both present different results based on the strength of correlations of structural predictors.
-According to \cite{Yehetal2014}, both solvent accessibility and  local packing density are predictors of site-wise variability with local packing density being strongest predictor of site-wise variability.
+According to \cite{Yehetal2014a}, both solvent accessibility and local packing density are predictors of site-wise variability with local packing density being strongest predictor of site-wise variability.
 However, recently, \cite{Shahmoradietal2014} compared local packing density, residue flexibility, and solvent accessibility in viral proteins and found that relative solvent accessibility is a stronger predictor of site-wise variability.
@@
 
-Proteins are subject to a number of biophysical and functional constraints \citep{Scherreretal2012, Wilkeetal2010}.
+The effect of protein-protein interactions(Fraser 2002)
+
+Proteins are subject to a number of biophysical and functional constraints \citep{Scherreretal2012, Wilkeetal2010, Sikoseketal2014, Liberlesetal2009, Huangetal2014, Fraseretal2002, Liaoetal2010, Serohijosetal2012}.
 These constraints result in site-specific patterns of sequence variability within a protein.
@@
 \\
-\indent Recently, \cite{Yehetal2014} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by Contact Number and Weighted Contact Number \citep{Liaoetal2005, Yehetal2014, Huangetal2014}, was the strongest determinant of site variability.
+\indent Recently, \cite{Yehetal2014a} performed a similar analysis on a series of enzyme monomer proteins and found that packing density, as defined by Contact Number and Weighted Contact Number \citep{Liaoetal2005, Yehetal2014a, Huangetal2014}, was the strongest determinant of site variability.
 Soon Afterwards, \cite{Shahmoradietal2014} also performed a site-wise analysis on a series of viral proteins.
 In this analysis they found that RSA had a slightly stronger correlation with site variability as opposed to local packing density.
-Moreover, the effect seen between CN and WCN was of a much smaller magnitude as compared to \cite{Yehetal2014}.
+Moreover, the effect seen between CN and WCN was of a much smaller magnitude as compared to \cite{Yehetal2014a}.
 It is not well understood the reasoning behind the difference in magnitude in the correlations seen between the two studies.
@@
     The results presented in this work are based on two datasets.
-The first is a dataset of $209$ monomeric enzymes taken from \cite{Huangetal2014}, originally from \cite{Yehetal2014}.The original dataset was comprised of $213$ proteins but we removed four of the proteins (PDB IDS: 1BBS, 1BS0, 1DIN, 2HPL) that had did not have data at insertion sites.
+The first is a dataset of $209$ monomeric enzymes taken from \cite{Huangetal2014}, originally from \cite{Yehetal2014a}.The original dataset was comprised of $213$ proteins but we removed four of the proteins (PDB IDS: 1BBS, 1BS0, 1DIN, 2HPL) that had did not have data at insertion sites.
 Briefly, these proteins are all enzyme monomers   randomly picked from the Catalytic Site Atlas $2.2.11$ \citep{Porteretal2004} with protein sizes in the sample ranging from $95$ to $1287$ residues in length.
@@
 For the purposes of our comparison between the two datasets of interest we used the CN and WCN for the enzyme proteins calculated in \cite{Huangetal2014} and the CN and WCN values for the viral proteins from \cite{Shahmoradietal2014}.
-In both of these works and in \cite{Yehetal2014}, WCN and CN are both defined the same.
+In both of these works and in \cite{Yehetal2014a}, WCN and CN are both defined the same.
 Contact Number is defined as the number of $C_{\alpha}$ within a pre-redefined radius, $r_0$.
@@
 
-	where $r_j$ is the length between the $C_{\alpha}$ of residue $i$ and residue $j$ in a protein of length $N$ \citep{Yehetal2014}.
+	where $r_j$ is the length between the $C_{\alpha}$ of residue $i$ and residue $j$ in a protein of length $N$ \citep{Yehetal2014a}.
 
@@
 
-We calculate the strength between the structural properties and site variability by calculating the Spearman correlation, $\rho$,  between the structural property (solvent exposure and local packing density) and site variability, as measured by either evolutionary rate or site entropy.
+Although it is clear that structure plays a role in the evolution of proteins, it is not clear to what extent protein structure constrains sequence variability.
+To date there have been conflicting reports about the strength of structure-sequence correlations \citep{Yehetal2014b, Shahmoradietal2014}.
+Therefore we sought to determine the main determinants of structure-sequence correlations within proteins.
+First, we calculated the strength between the structural properties and site variability by calculating the Spearman correlation, $\rho$,  between the structural property (solvent exposure and local packing density) and site variability, as measured by either evolutionary rate or site entropy.
 On average the correlations between CN and entropy are larger in absolute magnitude in the enzyme proteins as compared to viral proteins.
@@
 Figure 1 shows the correlations between CN and entropy for the viral and enzyme proteins.
-Even though on average the correlations in the enzyme proteins are larger as reported by \cite{Yehetal2014}, the viral proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations.
+Even though on average the correlations in the enzyme proteins are larger as reported by \cite{Yehetal2014a}, the viral proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations.
 
@@
 This means that enzyme and virus proteins have different correlations when using mean entropy as a predictor.
-This agrees with differences seen in the previous works by \cite{Yehetal2014} and \cite{Shahmoradietal2014}.
+This agrees with differences seen in the previous works by \cite{Yehetal2014a} and \cite{Shahmoradietal2014}.
 
@@
 \label{sec:dcr}
-Structure imposes constraints on the evolution of proteins.
+
+There has been previous work that has illuminated several factors that determine the rate at which proteins evolve \citep{Fraseretal2002, Drummondetal2008, Wilkeetal2010, Yangetal2012}.
+Some of these factors include: expression level \citep{Drummondetal2008, Wilkeetal2010, Paletal2001, Subramanianetal2004}, interactions with other protein partners \citep{Fraseretal2002, Yangetal2012, Minterisetal2005, Pangetal2010}  and selection for the costs of misfolding \citep{Drummondetal2005}.
+
+\indent Proteins are under a number of constraints for function \citep{Liberlesetal2009}.
+Biophysical considerations that have been shown to affect function include: selection for protein stability and proper folding \citep{Drummondetal2005}, specificity for protein binding \citep{Zarrinparetal2003}  and selection against non-specific and extraneous binding \citep{Levyetal2012}, selection to minimize protein aggregation \citep{DePristoetal2005}  and protein structure \citep{Fransozaetal2009, Fransozaetal2012, Yehetal2014a, Yehetal2014b, Shahmoradietal2014, Echaveetal2015, Huangetal2014}.
+Recently there have been several works that have looked into the effect of protein structure on protein function.
+Local packing density and solvent accessibility have been emerged as two structural predictors that have been showed to predict site variability \citep{Yehetal2014a, Yehetal2014b, Shahmoradietal2014, Fransozaetal2009, Bustamanteetal2000}.
+
+\indent Structure imposes constraints on the evolution of proteins.
 Therefore structural predictors can be used to predict the variability at sites.
@@
 Sites that are densely packed and have more contacts tend to evolve slower and exhibit less sequence variability.
+However, the strength of how well these two structural quantities (local packing density and solvent accessibility) has not been clearly explained.
+\cite{Yehetal2014a} examined the effect of using solvent accessibility and local packing density as measures of predicting site-wise variability.
+They found that both were predictors.
+However, \cite{Shahmoradietal2014} performed the same analysis on viral proteins and found that will both were predictors of stir-wise variability, the correlations between these two structural predictors and site variability were significantly lower.
 
-\indent Here we examined the relationship between site variability and the strength of structure-sequence relationships.
+\indent Here we examined the relationship between site variability and the strength of structure-sequence relationships in order to dissect the nature of the correlation between structural properties and site variability.
 We found that proteins with a larger variance in site variability as measured by sequence entropy, have stronger structure-sequence relationships on average.
 However, when looking at the mean entropy values of proteins there is a distinct difference between the virus and enzyme datasets.
-Overall the enzyme proteins have more site variability than the viral proteins.When predicting structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variability.
-Additionally the reason that the enzyme proteins in \cite{Yehetal2014} had higher correlation coefficients is that on average, the enzyme proteins have larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}.
+Overall the enzyme proteins have more site variability than the viral proteins.
+When predicting structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variability.
+\indent Additionally the reason that the enzyme proteins in \cite{Yehetal2014a} had higher correlation coefficients is that on average, the enzyme proteins have larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}.
 Moreover, since the variance in entropy is a more significant predictor of correlations in local packing density, an increased level of variance in site variability will have a larger effect on the local packing density correlations.
@@
 Therefore when examining structural determinants of site variability, entropy is a may be a more meaningful quantity.
-
-Maybe change the focus: forget trying to figure out which is the best predictor: RSA or local packing density.
-Focus on bridging the gap on why there are differences in correlation strength.
-They had stronger correlations in the Yeh et al (2014) paper as compared to the Shahmoradi paper.
-Make a table of the mean correlations and variance from both papers.
-
-Proteins are under a number of constraints for function (Liberles et al Paper).
-Biophysical considerations that have been shown to affect function include: selection for protein stability, specificity for protein binding and selection against extraneous binding.
-Recently there have been several works that have looked into the effect of protein structure on protein function as well (Xia, Yeh, Shahmoradi, Echave, Huang)
-
-{\color{red} Amir's comment: I think it maybe better to not put so much emphasis on variance of entropy, but its implications and meaning.
-For example entropy variance is an indicator of sequence divergence, to some extent.}
 
@@
 Proteins with a larger variance in site entropy among sites tend to smaller CN-Entropy correlations in magnitude.
-}
+Note: The entropy values in \cite{Shahmoradietal2014} were scaled.}
             \label{fig:entropy_cn_cor}
@@
 
+Make a table of the mean correlations and variance from both papers.
+
 \end{document}

### 2015-06-10 manuscript/manuscript.tex
Making final first edits

@@
 However, the strength and significance of these structure-sequence relations appear to vary widely among different proteins, with absolute correlation strengths ranging from $0.1$ to $0.8$.
-There have two works that have investigated structural predictors of site variability which both present different results based on the strength of correlations of structural predictors.
+Recently two works that have investigated structural predictors of site variability which both present different results based on the strength of correlations of structural predictors.
 According to \cite{Yehetal2014a}, both solvent accessibility and local packing density are predictors of site-wise variability with local packing density being strongest predictor of site-wise variability.
@@
 
-The effect of protein-protein interactions(Fraser 2002)
-
 Proteins are subject to a number of biophysical and functional constraints \citep{Scherreretal2012, Wilkeetal2010, Sikoseketal2014, Liberlesetal2009, Huangetal2014, Fraseretal2002, Liaoetal2010, Serohijosetal2012}.
-These constraints result in site-specific patterns of sequence variability within a protein.
+These constraints result in both global patterns between proteins and site-specific patterns of sequence variability within proteins.
 Recently several site-specific structural properties that can explain patterns of sequence variability in proteins have been identified.
@@
 To date there have been conflicting reports about the strength of structure-sequence correlations \citep{Yehetal2014b, Shahmoradietal2014}.
-Therefore we sought to determine the main determinants of structure-sequence correlations within proteins.
+Therefore we sought to determine the main determinants of the strength of structure-sequence correlations within proteins.
 First, we calculated the strength between the structural properties and site variability by calculating the Spearman correlation, $\rho$, between the structural property (solvent exposure and local packing density) and site variability, as measured by either evolutionary rate or site entropy.
-On average the correlations between CN and entropy are larger in absolute magnitude in the enzyme proteins as compared to viral proteins.
-The average $\rho$ between each structural property and each measure of site variability (entropy and evolutionary rates) can be seen in Table \ref{table:entropy_stats}  and Table \ref{table:rate_stats}.
-The viral proteins experience much lower site variability.
-Figure 1 shows the correlations between CN and entropy for the viral and enzyme proteins.
-Even though on average the correlations in the enzyme proteins are larger as reported by \cite{Yehetal2014a}, the viral proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations.
+The magnitude of these correlations allows us to determine the strength of solvent accessibility-site variability and local packing density - site variability correlations.
+For our analysis we used two measures of site variability, Shannon entropy and evolutionary rate.
+We calculated the LPD-site variability correlations for two types of local packing density: contact number (CN) and weighted contact number.
+We calculated the correlations for two sets of proteins.
+One was a set of monomeric enzyme proteins and the other was a set of virus proteins.
 
-\indent For a given protein its mean entropy and the variance might be different.
+\indent On average the correlations between CN and entropy are larger in absolute magnitude in the enzyme proteins as compared to viral proteins.
+The average $\rho$ between each structural property and each measure of site variability (entropy and evolutionary rates) can be seen in Tables \ref{table:entropy_stats} and \ref{table:rate_stats}.
+Both CN and WCN are positively correlated with  both measures of site variability.
+Residues that are more tightly packed (i.e., have more contacting neighbors) are more constrained and less variable.
+Solvent accessibility, as measured by relative solvent accessibility (RSA) is negatively correlated with site variability.
+Residues that are exposed on the surface of the protein are less constrained and exhibit more site variability.
+Both of these trends have been noted previously \citep{Yehetal2014a, Yehetal2014b, Shahmoradietal2014, Liaoetal2005, Fransozaetal2009}.
+Although similar in sign, the correlations between all structure-sequence relationships are smaller in magnitude for the virus proteins (Table \ref{table:rate_stats}).
+However, even though on average the correlations in the enzyme proteins are larger as reported by \cite{Yehetal2014a}, the viral proteins still have correlation strengths that are comparable to some of the enzyme proteins with lower correlations (Figures \ref{fig:entropy_cn_cor}, \ref{fig:entropy_wcn_cor}, \ref{fig:entropy_rsa_cor}).
+
+\indent As noted earlier, the virus proteins have lower levels of site variability (Figures \ref{fig:entropy_cn_cor}, \ref{fig:entropy_wcn_cor}, \ref{fig:entropy_rsa_cor}).
+Due to this fact, we investigated the effect of site variability on the structure-sequence correlations since the viral proteins have lower correlations overall as well.
+When examining the effect of site variability on the strength of structure-sequence correlations, we looked at both the variance of site variability and the mean of site variability.
+For a given protein its mean entropy and the variance might be different and both of these quantities measure different sequence variably properties.
 A protein can have a high mean entropy but have a low variance and vice versa.
 Figure \ref{fig:protein_ex_plot}A details the relationship between the average entropy across sites within proteins and the variance of entropy across sites.
-Additionally, the distribution of variance varies greatly between proteins even when they from the same dataset (Figure \ref{fig:protein_ex_plot}B).
+Additionally, the distribution of variance varies greatly between proteins even when they are from the same dataset (Figure \ref{fig:protein_ex_plot}B).
 Therefore the mean entropy of a protein as well as the variance in entropy at sites may be predictive for structure-sequence relationships.
@@
 However, when examining the variance of entropy there is a clear trend within the enzyme proteins.
-Proteins with a higher variance in site variability across the protein typically higher  correlations in magnitude (Figures \ref{fig:entropy_cn_cor}B, \ref{fig:entropy_wcn_cor}B, \ref{fig:entropy_rsa_cor}B).
-If you extrapolate the trend from the enzyme proteins, the viral proteins follow a similar trend.
-The correlation between the $\rho$ between RSA and entropy and the variance of entropy is positive.
+Proteins with a higher variance in site variability across the protein typically have higher  correlations in magnitude (Figures \ref{fig:entropy_cn_cor}B, \ref{fig:entropy_wcn_cor}B, \ref{fig:entropy_rsa_cor}B).The correlation between the $\rho$ between RSA and entropy and the variance of entropy is positive.
 Proteins with a larger variance in site entropy have the strongest correlations.
-The correlation between local packing density (both WCN and CN) and the variance of entropy is negative.
+The correlation between the $\rho$ between local packing density (both WCN and CN) and entropy and the variance of entropy is negative.
+Proteins with more variance in entropy have lower correlations in magnitude.
 Unlike entropy, there is no relationship between the variance of evolutionary rates at sites and any of the measured structural properties (Figures \ref{fig:var_rate_cor}, \ref{fig:var_rate_rsa_cor}).
 There also is a wide spread in the variance of the evolutionary rates across proteins for both the enzyme and viral proteins.
+If you extrapolate the trend from the enzyme proteins, the viral proteins follow a similar trend.
 
-\indent Although both evolutionary rate, as measured by Rate4Site and entropy are measures of site variability, these quantities are distinctly different.
+\indent Although both evolutionary rate, as measured by Rate4Site, and entropy are measures of site variability, these quantities are distinctly different.
 Rate4Site measures the rate at which a site changes over time whereas site entropy measure the absolute variation at a site.
@@
 It appears that the rate at which a site evolves is not important for predicting the strength of structure-sequence relationships.
-However, the absolute amount of variation at a site can be used to predict the strength of structure-sequence relationships.
+There is no relationship between the variance of evolutionary rate and any of the structure-sequence relationships (Figures \ref{fig:var_rate_cor}, \ref{fig:var_rate_rsa_cor}).
+However, the absolute amount of variation at a site, as measured by mean entropy, can be used to predict the strength of structure-sequence relationships.
 
@@
 When looking at the variance of entropy, proteins within both datasets at in a similar fashion and overall trends seen between the variance of site variation and the strength of structure-sequence relationships is preserved across both datasets.
+In summary, site variability appears to be the main determinant of protein structure-sequence relationships with the variance in entropy at a site as the best predictor of magnitude.
 
@@
 
-There has been previous work that has illuminated several factors that determine the rate at which proteins evolve \citep{Fraseretal2002, Drummondetal2008, Wilkeetal2010, Yangetal2012}.
+There has been previous work that has illuminated several factors that determine the rate at which proteins evolve.
 Some of these factors include: expression level \citep{Drummondetal2008, Wilkeetal2010, Paletal2001, Subramanianetal2004}, interactions with other protein partners \citep{Fraseretal2002, Yangetal2012, Minterisetal2005, Pangetal2010} and selection for the costs of misfolding \citep{Drummondetal2005}.
+Recently the effect of biophysical constraints on sequence variability have been considered.
+Biophysical considerations that have been shown to affect function include: selection for protein stability and proper folding \citep{Drummondetal2005}, selection for protein binding specifity \citep{Zarrinparetal2003}  and selection against non-specific and extraneous binding \citep{Levyetal2012}, selection to minimize protein aggregation \citep{DePristoetal2005} and protein structure \citep{Fransozaetal2009, Fransozaetal2012, Yehetal2014a, Yehetal2014b, Shahmoradietal2014, Echaveetal2015, Huangetal2014}.
 
-\indent Proteins are under a number of constraints for function \citep{Liberlesetal2009}.
-Biophysical considerations that have been shown to affect function include: selection for protein stability and proper folding \citep{Drummondetal2005}, specificity for protein binding \citep{Zarrinparetal2003}  and selection against non-specific and extraneous binding \citep{Levyetal2012}, selection to minimize protein aggregation \citep{DePristoetal2005}  and protein structure \citep{Fransozaetal2009, Fransozaetal2012, Yehetal2014a, Yehetal2014b, Shahmoradietal2014, Echaveetal2015, Huangetal2014}.
-Recently there have been several works that have looked into the effect of protein structure on protein function.
+\indent There have been several works that have looked into the effect of protein structure on protein function.
 Local packing density and solvent accessibility have been emerged as two structural predictors that have been showed to predict site variability \citep{Yehetal2014a, Yehetal2014b, Shahmoradietal2014, Fransozaetal2009, Bustamanteetal2000}.
-
-\indent Structure imposes constraints on the evolution of proteins.
-Therefore structural predictors can be used to predict the variability at sites.
-Two main determinants of variability at protein sites are solvent accessibility and local packing density.
 Sites that are on surface of the protein tend to have higher solvent accessibility and exhibit more site variability.
 Sites that are densely packed and have more contacts tend to evolve slower and exhibit less sequence variability.
-However, the strength of how well these two structural quantities (local packing density and solvent accessibility) has not been clearly explained.
+However, the strength of how well these two structural quantities (local packing density and solvent accessibility) correlate with variability at sites has not been clearly explained.
 \cite{Yehetal2014a} examined the effect of using solvent accessibility and local packing density as measures of predicting site-wise variability.
-They found that both were predictors.
-However, \cite{Shahmoradietal2014} performed the same analysis on viral proteins and found that will both were predictors of stir-wise variability, the correlations between these two structural predictors and site variability were significantly lower.
+They found that both were significant predictors.
+However, \cite{Shahmoradietal2014} performed the same analysis on viral proteins and found that while both were predictors of stie-wise variability, the correlations between these two structural predictors and site variability were significantly lower.
 
 \indent Here we examined the relationship between site variability and the strength of structure-sequence relationships in order to dissect the nature of the correlation between structural properties and site variability.
-We found that proteins with a larger variance in site variability as measured by sequence entropy, have stronger structure-sequence relationships on average.
-However, when looking at the mean entropy values of proteins there is a distinct difference between the virus and enzyme datasets.
+We used both enzyme proteins and virus proteins in our analysis.
+When looking at the mean entropy values of proteins there is a distinct difference between the virus and enzyme datasets.
 Overall the enzyme proteins have more site variability than the viral proteins.
-When predicting structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variability.
-\indent Additionally the reason that the enzyme proteins in \cite{Yehetal2014a} had higher correlation coefficients is that on average, the enzyme proteins have larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}.
-Moreover, since the variance in entropy is a more significant predictor of correlations in local packing density, an increased level of variance in site variability will have a larger effect on the local packing density correlations.
-It appears that site entropy and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships.
-Finally, the variance in evolutionary rate among sites of a protein has a smaller an effect on the correlation magnitude as compared to entropy.
+We found that proteins with a larger variance in site variability as measured by sequence entropy, have stronger structure-sequence relationships on average and that when predicting structure-sequence relationships the variance in site variability among proteins is a better predictor than mean site variability.
+Both \cite{Yehetal2014a} and \cite{Shahmoradietal2014} report different levels of magnitude between LPD and RSA and site variability.
+According to our analysis, one reason that the enzyme proteins in \cite{Yehetal2014a} had higher correlation coefficients is that, on average, the enzyme proteins have a larger variance in entropy as compared to the virus proteins of \cite{Shahmoradietal2014}.
+
+\indent For all proteins, It appears that site entropy and evolutionary rate are not equal measures of site variability in terms of predicting the strength of structure-sequence relationships.
+The variance in evolutionary rate among sites of a protein has a smaller an effect on the correlation magnitude as compared to entropy.
 This suggests that the absolute variability at sites is a better predictor of whether structural determinants can be used to predict variability at sites  as opposed to evolutionary rate.
-Therefore when examining structural determinants of site variability, entropy is a may be a more meaningful quantity.
+Therefore when examining structural determinants of site variability, entropy may be a more meaningful quantity.
+The variance in entropy of a protein measures the variation in site variability among sites within a protein.
+The variance in entropy is not necessarily related to high sequence divergence.
+If divergence is high for a given protein but all sites are have equal levels of divergence, then the variance in entropy for a protein will be low.
+The same can be said for proteins with sites that are equal in terms of having a low level of variability among sites.
+If a protein has sites that are extremely variable (ex.
+some sites that are highly conserved and some that are less conserved), then they will have an have a high variance in entropy.
+A larger variation in variability within a protein allows a larger range of correlations with structural quantities.
+If all sites have similar levels of divergence (ex.
+low divergence or high divergence) then the dynamic range of correlations you expect to see with structural properties is limited.
+
+\indent We performed our analysis on two type of protein datasets, viral protein and enzyme proteins in order to verify the generality of our analysis.
+Both proteins exhibit some of the same selective pressures such as selection for stability and pressure to fold and adopt the correct native conformation.
+However, these proteins are very different in terms of function.
+Enzymes are used to catalyze chemical actions and as such have additional constraints such as structural constraints for a proper active site for catalytic function.
+On the other hand, viruses use their protein to infect and replicate within their hosts.
+These proteins are utilized to perform a variety of necessary functions for viral replication such as host cellular entry \citep{Radoshitskyetal2007, Allisonetal2014} and nuclear importation \citep{Schalleretal2014}.
+As host immune systems attack these viruses, these viruses evolve to escape from these host mechanisms resulting in signatures of positive selection within these proteins.
+There are several examples of positive selection seen in viral proteins (for a review see: \cite{Sironietal2015}).
+
+\indent Because of the difference for selective pressures for function facing these protein types there could different constraints on sequence variability.
+Indeed, we find that on average the enzyme proteins have more site variability and that virus proteins have a smaller spread in site variability.
+Most of the viral proteins analyzed have very low variably while the enzyme dataset has proteins with low-range, mid-range and high-range proteins in terms of site variability.
+Additionally the enzyme proteins are also more variability in terms of site variability of site within a protein as evidence by the variance in entropy.
+These results there are structural additional constraints that are different within viral proteins and enzyme proteins.
+Indeed recently \cite{Meyeretal2015} found that a geometrical model that includes distance to the receptor-binding region in Influenza A H3 hemmagglutinin (HA) was able to explain some of the the variance seen in evolutionary rate between sites within HA.
+This suggests that residues close to the host cell receptor-binding region experience different evolutionary constraints compared to residues farther away from the receptor-binding region.
+It is possible that a similar  distance metric for enzymes might be effective for understanding the structural constraints of selection for active site.
+Residues close to the active site may experience constraints in the same way.
+Therefore although general biophysical properties have been demonstrated to pertain all proteins, there are additional structural constraints that effect site-wise variably within different types of proteins.
+Further research into the specific biophysical constraints of difference classes of proteins (ex.
+viral, enzymes) will help better our understanding of how protein biophysics affects the evolution of proteins.
 
@@
 Proteins with a larger variance in site entropy among sites tend to smaller CN-Entropy correlations in magnitude.
-Note: The entropy values in \cite{Shahmoradietal2014} were scaled.}
+Note: The entropy values in this manuscript are different than in \cite{Shahmoradietal2014}.
+Those entropy values were scales by a a factor of $1/\sum_j P_{ij}$.
+}
             \label{fig:entropy_cn_cor}
@@
 
-Make a table of the mean correlations and variance from both papers.
-
 \end{document}
